### 2017-02-01 analysis/miss_position.Rmd
Files commited by wflow_commit.

---
title: "corss validation of missing value"
author: "wei wang"
date: 2017-01-30
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## 5-fold CV

start from five fold and can easily extend to 10 fold or K -fold.

### code

To generate the cross validation data for missing value, I randomly pick the missing possition for each fold.

the code is like the following:

```{r}

cross_plot = function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  missing= array(0,dim = c(k_fold,N,P))
  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  par(mfrow = c(2,3), mar = (c(5,4,4,2)-1.9))
  image(Y,main = "data Y")
  for(i in 1:k_fold){
    image(missing[i,,],main = "missing data")
  }
  return(missing)
}

```

### block design

Starting from simplest case, where Y is $5 \times 5$ matrix.

```{r}
N = 5
P = 5
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```

### permute rows and columns

for larger matrix

```{r}
N = 50
P = 100
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```
## Session Information

```{r session-info}
```