### 2017-02-02 analysis/miss_position.Rmd
Files commited by wflow_commit.

---
title: "corss validation of missing value"
author: "wei wang"
date: 2017-01-30
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## 3-fold CV

take 3-fold CV as example. We want to divide the data $Y$ in to 3 parts to do 3-fold CV. We randomly divide the the columns into 3 sets and the rows into 3 sets as well. The data matrix $Y$ is divided into 9 partition (by row and column permutation):

$$Y = \begin{pmatrix} Y_{11} & Y_{12} & Y_{13} \\ Y_{21} & Y_{22} & Y_{23} \\ Y_{31} & Y_{32} & Y_{33} \end{pmatrix}$$

Then $Y_{(1)} = \{  Y_{11},  Y_{22} , Y_{33}\}$, $Y_{(2)} = \{  Y_{12},  Y_{23} , Y_{31}\}$ and $Y_{(3)} = \{  Y_{13},  Y_{21} , Y_{32}\}$ are orthogonal to each other. Then the data matrix $Y$ is marked as:
$$Y = \begin{pmatrix} Y_{(1)} & Y_{(2)} & Y_{(3)} \\ Y_{(3)} & Y_{(1)} & Y_{(2)} \\ Y_{(2)} & Y_{(3)} & Y_{(1)} \end{pmatrix}$$

For complete data set, we would use scheme 1. The imputation methods we consider are based on the model with low rank structure: $LF^T$, so each $Y_{(i)}$ in scheme 1 contains equally one set of indices of $L_{.k}$ and $F_{k.}$. This also helps to make sure that all $i$'s and $j$'s are included into each $Y_{-(i)}$. Otherwise, we can't make prediction when some times whole columns or rows are missing.

$$Y=  \begin{pmatrix} Y_{11} & Y_{12} \\ Y_{21} & Y_{22}  \end{pmatrix}= \begin{bmatrix} L^{(1)}F^{(1)} & L^{(1)}F^{(2)} \\ L^{(2)}F^{(1)} & L^{(2)}F^{(2)} \end{bmatrix} + E = \left[ \begin{array}{c} L^{(1)} \\ L^{(2)} \end{array} \right] \times \left[ \begin{array}{cc} F^{(1)}& F^{(2)} \end{array} \right] + E$$

In this case, $Y_{(1)} = \{  Y_{11},  Y_{22} \}$, $Y_{(2)} = \{  Y_{12},  Y_{21} \}$, and each of $L^{(1)},F^{(1)},L^{(2)},F^{(2)}$ shows up in both $Y_{(1)}$ and $Y_{(2)}$ so the information contains in each $Y_{(i)}$ are in balance.

## 5-fold CV

start from five fold and can easily extend to 10 fold or K -fold.

### code

To generate the cross validation data for missing value, I randomly pick the missing possition for each fold.

the code is like the following:

```{r}

cross_plot = function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  missing= array(0,dim = c(k_fold,N,P))
  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  par(mfrow = c(2,3), mar = (c(5,4,4,2)-1.9))
  image(Y,main = "data Y")
  for(i in 1:k_fold){
    image(missing[i,,],main = "missing data")
  }
  return(missing)
}

```

### block design

Starting from simplest case, where Y is $5 \times 5$ matrix.

```{r}
N = 5
P = 5
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```

### permute rows and columns

for larger matrix

```{r}
N = 50
P = 100
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```
## Session Information

```{r session-info}
```