### 2017-02-13 analysis/sd_chr1_asinh.Rmd
Files commited by wflow_commit.

---
title: "FLASH on chrommarks_chipcounts_example_regions"
author: "wei wang"
date: 2017-02-09
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}

#GM12878 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/GM12878_10marks_meanChIP_chr1_region_asinh.rds")
#Y = GM12878[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_GM = b_flash
#saveRDS(FL_GM, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_GM_var_col.rds")
FL_GM = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_GM_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}

HUVEC = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/HUVEC_10marks_meanChIP_chr1_region_asinh.rds")
#Y = HUVEC[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_HUVEC = b_flash
#saveRDS(FL_HUVEC, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_HUVEC_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
#K562 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/K562_10marks_meanChIP_chr1_region_asinh.rds")
#Y = K562[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_K562 = b_flash
#saveRDS(FL_K562, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_K562_var_col.rds")
FL_K562 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_K562_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
#NHLF = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/NHLF_10marks_meanChIP_chr1_region_asinh.rds")
#Y = NHLF[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_NHLF = b_flash
#saveRDS(FL_NHLF, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_NHLF_var_col.rds")
FL_NHLF = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_NHLF_var_col.rds")
```

## factor plot

### GM12878

```{r,echo=FALSE}
par(mfrow = c(2,2))
for(i in 1:4) barplot(FL_GM$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,":GM12878"),las = 2,cex.names = 0.6)
```

### K562

```{r,echo=FALSE}
par(mfrow = c(2,3))
for(i in 1:5) barplot(FL_K562$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,"K562"),las = 2,cex.names = 0.6)
```

### NHLF

```{r,echo=FALSE}
par(mfrow = c(2,3))
for(i in 1:5) barplot(FL_NHLF$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,"NHLF"),las = 2,cex.names = 0.6)
```

## Session Information

```{r session-info}
```