### 2017-03-13 analysis/simulationrank1.Rmd
Files commited by wflow_commit.

---
title: "simulation for rankone case"
author: "Wei Wang"
date: 2017-03-13
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## simulation rank one

### simulated data

```{r,echo=FALSE}
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true))
}
# I think we can try some situation
# pve is small due to the

# this can be a example of the bi sparse factor and loadings

CVPMD=function(Y,c_u,c_v){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_u = length(c_u)
  n_v = length(c_v)
  CVRMSE = array(0,dim = c(n_u,n_v))
  minrmse = Inf
  opt_u = 0
  opt_v = 0
  for(t_u in 1:n_u){
    for(t_v in 1:n_v){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v])
        misshat = out$d * out$u %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_u,t_v] = mean(rmse)
      if(CVRMSE[t_u,t_v] < minrmse){
        minrmse = CVRMSE[t_u,t_v]
        opt_u = c_u[t_u]
        opt_v = c_v[t_v]
      }
    }
  }
  return(list(opt_u = opt_u,opt_v = opt_v))
}

PMA.wrapper = function(Y,ngrids = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_u = seq(1,sqrt(N),length.out = ngrids)
  c_v = seq(1,sqrt(P),length.out = ngrids)
  cvout = CVPMD(Y,c_u,c_v)
  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v )
  residual_PMD = Y - out$d * out$u %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}
```

## big signal

```{r}
set.seed(99)
Data = datamaker(60,100,c(0.5,0.1,0.2,0.15,0.05),c(0.01,0.5,1,2,5),0.5,c(0.6,0.05,0.3,0.025,0.025),c(0.01,0.5,1,2,5),0.5,sqrt(1))
dim(Data$Y)
gf = flashr::flash(Data$Y)
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
abline(0,1,col = "red")
PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
```

```{r}
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
abline(0,1,col = "red")
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
sqrt(mean((gf$l %*% t(gf$f) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
sqrt(mean(((gPMD$u * gPMD$d) %*% t(gPMD$v) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
plot(as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "blue")
```

## Session Information

```{r session-info}
```