### 2017-03-13 analysis/simulation2.Rmd
Files commited by wflow_commit.

---
title: "Simulation with rank 2"
author: "Wei Wang"
date: 2017-03-13
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## simulated data

```{r}
set.seed(99)
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}
N = 180
P = 300
K = 2
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[,1] = rnorm(P,0,1)
F_true[,2] = rnorm(P,0,1)
# F_true = scale(F_true)
rho = rnorm(N,1.96,0.6)
L_true[,1] = rho * sin(pi*(seq(1,N)/(N/2)))
L_true[,2] = rho * cos(pi*(seq(1,N)/(N/2)))
group1_index = c(31:60,121:150)
group2_index = c(1:15,76:105,166:180)
group3_index = seq(1:180)[-(union(group1_index,group2_index))]
# G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,1),nrow=N)
Y = L_true %*% t(F_true) + E
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(L_true[,1])
plot(L_true[,2])
g_flash = flashr::greedy(Y,K = 5)
gPMD = PMA::PMD(Y,K = 2)
gsvd = svd(Y,nv = 2,nu = 2)
```

```{r}
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(g_flash$l[,1],g_flash$l[,2],col = "green")
points(g_flash$l[group1_index,1],g_flash$l[group1_index,2],col = "green")
points(g_flash$l[group2_index,1],g_flash$l[group2_index,2],col = "red")
points(g_flash$l[group3_index,1],g_flash$l[group3_index,2],col = "blue")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(gPMD$u[,1],gPMD$u[,2],col = "green")
points(gPMD$u[group1_index,1],gPMD$u[group1_index,2],col = "green")
points(gPMD$u[group2_index,1],gPMD$u[group2_index,2],col = "red")
points(gPMD$u[group3_index,1],gPMD$u[group3_index,2],col = "blue")
plot(gsvd$u[,1],gsvd$u[,2],col = "green")
points(gsvd$u[group1_index,1],gsvd$u[group1_index,2],col = "green")
points(gsvd$u[group2_index,1],gsvd$u[group2_index,2],col = "red")
points(gsvd$u[group3_index,1],gsvd$u[group3_index,2],col = "blue")
```

## Session Information

```{r session-info}
```