### 2017-03-15 analysis/simulation2.Rmd
Files commited by wflow_commit.

---
title: "Simulation with rank 2"
author: "Wei Wang"
date: 2017-03-13
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

###
Here I would like to add a cv function for PMD when K larger than 1

In the PMA R package we can find that they use the parameter sumabs to control the sparsity, so we use this one as well.

```{r}
CVPMD=function(Y,c_s,K){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_s = length(c_s)
  CVRMSE = rep(NA,n_s)
  minrmse = Inf
  opt_s = 0
  # here we only need one layer
    for(t_s in 1:n_s){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_s] = mean(rmse)
      if(CVRMSE[t_s] < minrmse){
        minrmse = CVRMSE[t_s]
        opt_s = c_s[t_s]
      }
    }

  return(list(opt_s = opt_s ))
}

PMA.wrapper = function(Y,ngrids = 10,K){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.8,len=ngrids)
  cvout = CVPMD(Y,c_s, K = K)
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}
```

## simulated data

```{r}
set.seed(99)
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}
N = 100
P = 200
K = 2
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[,1] = rnorm(P,0,0.3)
F_true[,2] = rnorm(P,0,0.3)
# F_true = scale(F_true)
group1_index = c(1:30)
group2_index = c(71:100)
group3_index = seq(1:90)[-(union(group1_index,group2_index))]
L_true[group1_index,1] = betaO(30,c(0.9,0.1),c(0.001,0.1),1)
L_true[group2_index,1] = betaO(30,c(0.9,0.1),c(0.5,0.75),1)
L_true[group3_index,1] = betaO(40,c(0.9,0.1),c(1,1.5),1)
L_true[group1_index,2] = betaO(30,c(0.9,0.1),c(1,1.5),1)
L_true[group2_index,2] = betaO(30,c(0.9,0.1),c(0.5,0.75),1)
L_true[group3_index,2] = betaO(40,c(0.9,0.1),c(0.001,0.1),1)
# generate Y
E = matrix(rnorm(N*P,0,1),nrow=N)
Y = L_true %*% t(F_true) + E
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(L_true[,1])
plot(L_true[,2])
g_flash = flashr::greedy(Y,K = 5)
g_flash = flashr::backfitting(Y,initial_list = g_flash,maxiter_bf = 10)
gPMD = PMA.wrapper(Y,ngrids = 10,K = 2)
gsvd = svd(Y,nv = 2,nu = 2)
```

```{r}
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(g_flash$l[,1],g_flash$l[,2],col = "green")
points(g_flash$l[group1_index,1],g_flash$l[group1_index,2],col = "green")
points(g_flash$l[group2_index,1],g_flash$l[group2_index,2],col = "red")
points(g_flash$l[group3_index,1],g_flash$l[group3_index,2],col = "blue")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(gPMD$u[,1],gPMD$u[,2],col = "green")
points(gPMD$u[group1_index,1],gPMD$u[group1_index,2],col = "green")
points(gPMD$u[group2_index,1],gPMD$u[group2_index,2],col = "red")
points(gPMD$u[group3_index,1],gPMD$u[group3_index,2],col = "blue")
plot(gsvd$u[,1],gsvd$u[,2],col = "green")
points(gsvd$u[group1_index,1],gsvd$u[group1_index,2],col = "green")
points(gsvd$u[group2_index,1],gsvd$u[group2_index,2],col = "red")
points(gsvd$u[group3_index,1],gsvd$u[group3_index,2],col = "blue")
plot(gsvd$d)
```

```{r}
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(as.vector(L_true %*% t(F_true)),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "red")
abline(0,1,col = "red")
plot(as.vector(L_true %*% t(F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
abline(0,1,col = "red")
sqrt(mean((g_flash$l %*% t(g_flash$f) - L_true %*% t(F_true))^2))/sqrt(mean((L_true %*% t(F_true))^2))
sqrt(mean((gPMD$u %*% diag(gPMD$d) %*% t(gPMD$v) - L_true %*% t(F_true))^2))/sqrt(mean((L_true %*% t(F_true))^2))
plot(as.vector( gPMD$u %*% diag(gPMD$d) %*% t(gPMD$v) ),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "blue")
abline(0,1,col = "red")
plot(as.vector( gsvd$u %*% diag(gsvd$d[1:2]) %*% t(gsvd$v) ),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "blue")
abline(0,1,col = "red")
```

## Session Information

```{r session-info}
```