### 2017-03-30 analysis/GTExresult.Rmd
Files commited by wflow_commit.

---
title: "result for GTEx data and"
author: "wei wang"
date: 2017-03-30
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

The result for FLASH on GTEx data is on midway cluster

## Zsocre data

To get the zscore data:

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/gtexEQTL_zscore.rds ./
```

and then open `Rstudio`

```
load("gtexEQTL_zscore.rds")
```

the matrix `zscore` is the data matrix.

## FLASH result

### constant variance

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashconst.rds ./
```

### column-wise variance

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashvarcol.rds ./
```

### constant variance + known variance structure

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashnoisyConst.rds ./
```

### column-wise variance + known variance structure

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashnoisycol.rds ./
```

### load the result

we load the column-wise variance reslut which we use in our paper as example

```
load("bflashvarcol.rds")
```

and the list `b_flash` is the reslut of backfitting FLASH with estimation of $l$ $f$ and $\hat{\Sigma}_e$

## plot the barplot

this is from Gao:

data set: /project/mstephens/data/internal_supp/gtex-v6-sumstat-hdf5

[5:25]  
on midway. you can focus on `MatrixEQTLSumStats.Portable.h5`

[5:26]  
for the 16000 times 44 data-set, stored in a table called `max` (edited)

[5:26]  
To process the data, https://github.com/stephenslab/gtex-eqtls/blob/master/doc/notes/SumstatsDB.md

[5:27]  
and the relevent scripts in this MD file: https://github.com/stephenslab/lab-resource/blob/master/utilities/src/SumstatQuery.R

[5:30]  
you should use this line `GetSS("max", "/project/mstephens/gtex/analysis/april2015/query/MatrixEQTLSumStats.h5")`

----- October 20th, 2016 -----
gaow [6:06 PM] 
https://github.com/stephenslab/lab-resource/wiki/Tissue-Color-Palette

[6:06]  
see "load to R"

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```