### 2017-04-11 analysis/simulation3.Rmd
Files commited by wflow_commit.

---
title: "simulation for the sparse factor loadings"
author: "Wei Wang"
date: 2017-04-04
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}

CVPMD=function(Y,c_s,K){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_s = length(c_s)
  CVRMSE = rep(NA,n_s)
  minrmse = Inf
  opt_s = 0
  # here we only need one layer
  for(t_s in 1:n_s){
    rmse = rep(0,5)
    for(i in 1:5){
      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
      misshat = out$u %*%  diag(out$d) %*% t(out$v)
      for(j in 1:5){
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
      }
      rmse[i] = sqrt(rmse[i] / (N * P/5))
    }
    CVRMSE[t_s] = mean(rmse)
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }

  return(list(opt_s = opt_s ))
}

PMA.wrapper = function(Y,ngrids = 10,K){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.8,len=ngrids)
  cvout = CVPMD(Y,c_s, K = K)
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}

```

```{r}
set.seed(99)
N = 150
P = 240
K = 3
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[1:80,1] = rnorm(80,0,0.5)
F_true[81:160,2] = rnorm(80,0,1)
F_true[161:240,3] = rnorm(80,0,2)
L_true[1:10,1] = rnorm(10,0,2)
L_true[11:60,2] = rnorm(50,0,1)
L_true[61:150,3] = rnorm(90,0,0.5)
G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,2),nrow=N)
Y = L_true %*% t(F_true) + E
g_flash = flashr::greedy(Y,K = 5)
g_b_flash = flashr::backfitting(Y,initial_list = g_flash)

outcv = PMA.wrapper(Y,10,K = 3)
```

```{r}
library(reshape2)
melted_cormat_1 <- melt(G, na.rm = TRUE)
melted_cormat_2 <- melt(g_b_flash$l %*% t(g_b_flash$f), na.rm = TRUE)
melted_cormat_3 <- melt(outcv$u %*% diag(outcv$d) %*% t(outcv$v), na.rm = TRUE)
melted_cormat_4 <- melt(Y, na.rm = TRUE)
# Heatmap
library(ggplot2)
p1 = ggplot(data = melted_cormat_1, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p2 = ggplot(data = melted_cormat_2, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "FLASH",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p3 = ggplot(data = melted_cormat_3, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "PMD",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p4 = ggplot(data = melted_cormat_4, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "observed data",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

library(cowplot)
plot_grid(p1,p4,p3,p2, ncol = 2)
```

```{r}
library(ggplot2)
library(cowplot)
mat = cbind(L_true,g_b_flash$l,outcv$u)
mat[,8] = -mat[,8]
mat[,9] = -mat[,9]
mat[,6] = -mat[,6]
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth.l1","truth.l2","truth.l3","FLASH.l3","FLASH.l2","FLASH.l1","PMD.u3","PMD.u2","PMD.u1")
df1 = data.frame(mat1)
p1_1_l <- ggplot(df1, aes(truth.l1, FLASH.l1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (FLASH)")
p1_2_l <- ggplot(df1, aes(truth.l2, FLASH.l2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (FLASH)")
p1_3_l <- ggplot(df1, aes(truth.l3, FLASH.l3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (FLASH)")
p2_1_l <- ggplot(df1, aes(truth.l1, PMD.u1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (PMD)")
p2_2_l <- ggplot(df1, aes(truth.l2, PMD.u2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (PMD)")
p2_3_l <- ggplot(df1, aes(truth.l3, PMD.u3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (PMD)")

library(ggplot2)
library(cowplot)
mat = cbind(F_true,g_b_flash$f,outcv$v)
mat[,8] = -mat[,8]
mat[,9] = -mat[,9]
mat[,6] = -mat[,6]
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth.f1","truth.f2","truth.f3","FLASH.f3","FLASH.f2","FLASH.f1","PMD.v3","PMD.v2","PMD.v1")
df1 = data.frame(mat1)
p1_1 <- ggplot(df1, aes(truth.f1, FLASH.f1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (FLASH)")
p1_2 <- ggplot(df1, aes(truth.f2, FLASH.f2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (FLASH)")
p1_3 <- ggplot(df1, aes(truth.f3, FLASH.f3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (FLASH)")
p2_1 <- ggplot(df1, aes(truth.f1, PMD.v1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (PMD)")
p2_2 <- ggplot(df1, aes(truth.f2, PMD.v2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (PMD)")
p2_3 <- ggplot(df1, aes(truth.f3, PMD.v3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (PMD)")
plot_grid(p1_1_l, p2_1_l, p1_1, p2_1,
          p1_2_l, p2_2_l, p1_2, p2_2,
          p1_3_l, p2_3_l, p1_3, p2_3,labels = "" , ncol = 4)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```