### 2017-04-13 analysis/DNAdata.Rmd
Files commited by wflow_commit.

---
title: "Untitled"
author: "First Last"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
library(readxl)
Data = read_excel("../data/DNAdata/sd03.xls")
data = as.matrix(Data[,-c(1,2)])
#saveRDS(data,file = "../data/DNAdata/sd03.rds")
```

```
source("~/HG/flash/Rcode/flash.R")
source("~/HG/flash/Rcode/greedy.R")
source("~/HG/flash/Rcode/backfitting.R")
Y = readRDS("sd03.rds")
Y = t(Y)
dim(Y)
g_flash = greedy(Y, K = 30, flash_para = list(partype = "var_col"))
saveRDS(g_flash, "./logcpm/gflash_var_col_imput.rds")

b_flash = backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 6)
saveRDS(b_flash, "./logcpm/bflash_var_col_imput.rds")
```

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.err
#SBATCH --array=1
#SBATCH --time=32:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=32G

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/DNAdata
Rscript --verbose /home/weidong/HG/flash/data/DNAdata/flash_const_wrapper.R

```

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.err
#SBATCH --array=1
#SBATCH --time=32:00:00
#SBATCH --partition=mstephens
#SBATCH --account=pi-mstephens
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --mem-per-cpu=32G

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/DNAdata
Rscript --verbose /home/weidong/HG/flash/data/DNAdata/flash_varcol_wrapper.R
```

```
gc = readRDS("../data/DNAdata/b_flash_const.rds")
gv = readRDS("../data/DNAdata/bflash_varcol.rds")
gsvd = svd(t(data))
library(flashr)
library(reshape2)
melted_data <- melt((gv$l %*% t(gv$f))[1:30,1:20000], na.rm = TRUE)
# Heatmap
library(ggplot2)
p1 = ggplot(data = melted_data, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "black", high = "red", mid = "blue", 
   midpoint = 0, limit = c(-0.6,0.6), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

image(t(gv$l %*% t(gv$f)),col = rainbow(7))

i = 3
par(mfrow = c(2,2))
barplot(gc$l[,i])
barplot(gv$l[,i])
barplot(gsvd$u[,i])
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```