### 2017-04-27 analysis/meetnotes.Rmd
Files commited by wflow_commit.

---
title: "Meeting notes"
author: "Wei"
date: 2017-04-27
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## meeting summary 04-26

### on flash paper

#### OCV criteria

we will introduce OCV in FLASH paper for missing data impute. But the criteria of the OCV is RMSE of the data.
$$RMSE(\hat{Y},\Omega) = \sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Y_{(ij)} - \hat{Y}_{(ij)})^2}$$
 where the $\Omega$ is the set of missing index. However we discussed that there is another criteria we can use.

 First fit the model 
 $$Y_{ij} = \alpha_i +\beta_j + N(0,\sigma_i^2 \sigma_j^2)$$
 $$Z_{ij} = \frac{Y_{ij} - \alpha_i -\beta_j}{\sigma_i \sigma_j} $$
and use

$$\frac{\sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Z_{(ij)} - \hat{Z}_{(ij)})^2}}{\sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Z_{(ij)} )^2}}$$

To compare which one is better, we will try both method and compare the result. In the second criteria, we can impute the missing data and transform $Z_{ij}$ back to $Y_{ij}$ to make the results comparable.

#### On the lemma

In our meeting, Matthew mentioned that we can provide a general lemma to show that the VEM factor analysis can be convert to EB approach of normal means problem.

#### psuedo algorithm

I should add the step of estimating $\sigma^2$ into the psuedo algorithm

#### combine the marginalization on $f$ and maximization on $l$ together

The two steps are basically for one problem so we can combine these two subsections together.

#### barplot for flash result

In the Gtex data, the plot is too big to fit one page. try to break the plot into two pages.

## Session Information

```{r session-info}
```