### 2017-05-01 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

---
title: "pairwise plot of flash result of flash Z-score"
author: "Wei Wang"
date: 2017-05-01
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
load("../data/GTExdata/bflashvarcol.rds")
dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
```

```{r}
plot(b_flash$l[,1],b_flash$l[,3],col = as.vector(colordata[,2]),pch = 16)
text(b_flash$l[,1],b_flash$l[,3], labels=colordata[,1], cex= 0.7)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```