### 2017-05-03 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

---
title: "pairwise plot of flash result of flash Z-score"
author: "Wei Wang"
date: 2017-05-01
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
load("../data/GTExdata/bflashvarcol.rds")
dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
```

### plot of $l_1$ vs $l_3$

```{r}
i = 1
j = 3
plot(b_flash$l[,i],b_flash$l[,j],col = as.vector(colordata[,2]),pch = 16)
text(b_flash$l[,i],b_flash$l[,j], labels=colordata[,1], cex= 0.7)
```

### plot $f_1$ vs $f_3$

```{r}
i = 1
j = 3
plot(b_flash$f[,i],b_flash$f[,j],col = "grey",pch = 16)

```

## logcpm of count

here I use
- 2 brain tissues
- testis + pituitary
- 2 skin
- Mammary+ adipose
- 2 heart
- uterus + Ovary

for each tissue, we random pick 50 smaples.

```{r}
bfl = readRDS("../data/GTExdata/logcpm/sixpairssubsample/bflash_varcol.rds")
```

I run flash with K=60 and it provide 60 factors in greedy algorithm, which mean it didn't stop. I run backfitting procedure with 6 iteration due to the time limit, it still provide 60 factors.

I am not sure the factor are useful a

```{r}
color_six = c(rep("#eeee00",100),rep("#aaaaaa",50),rep("#aaff99",50),
              rep("#0000ff",50),rep("#7777ff",50),
              rep("#ffaa00",50),rep("#33cccc",50),
              rep("#9900ff",50),rep("#660099",50),
              rep("#ff66ff",50),rep("#ffaaff",50))

par(mfrow = c(3,3),mar = c(5,4,4,2)-1.8)
for(i in 1:60){
  barplot(bfl$l[,i],col = color_six,border = color_six)
}
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```