### 2017-05-22 analysis/imputestratergy.Rmd
Files commited by wflow_commit.

---
title: "impute stragergy"
author: "Wei Wang"
date: 2017-05-22
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## try the breast cancer data

```{r}
library(R.matlab)
library(PMA)
datapath = "~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/nsfaexample.mat"
Y_centered = readMat(datapath)
Y = Y_centered$Ycentered
library(flashr)
library(softImpute)
## to get the biscale data
Y_bs = biScale(Y)

para_list = attributes(Y_bs)

set.seed(999)
# set the missing position

row_index = sample(dim(Y)[1],20)
col_index = sample(dim(Y)[2],20)
Y_miss = Y[row_index,col_index]

Y_input = Y
Y_input[row_index,col_index] = NA
Y_bs_input = Y_bs
Y_bs_input[row_index,col_index] = NA

#g1 = flashr::greedy(Y_input, K  = 60)
#g2 = flashr::greedy(Y_bs_input, K  = 60)
#saveRDS(g1,"../data/missingvalue/imputescheme/g1_original.rds")
#saveRDS(g2,"../data/missingvalue/imputescheme/g2_biscale.rds")
g1 = readRDS("../data/missingvalue/imputescheme/g1_original.rds")
g2 = readRDS("../data/missingvalue/imputescheme/g2_biscale.rds")
Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
sqrt(mean((Y_miss - Y_hat_1)^2))
Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
```

```
#### for RCC
library(R.matlab)
library(PMA)
datapath = "/home/weidong/HG/flash/data/missingvalue/Breastcancer/nsfaexample.mat"
Y_centered = readMat(datapath)
Y = Y_centered$Ycentered
source("~/HG/flash/Rcode/flash.R")
source("~/HG/flash/Rcode/greedy.R")
source("~/HG/flash/Rcode/backfitting.R")
library(softImpute)
## to get the biscale data
Y_bs = biScale(Y)
para_list = attributes(Y_bs)

set.seed(99)
results = matrix(0,ncol = 2, nrow = 30)
for(i in 1:30){
  print(i)
  row_index = sample(dim(Y)[1],20)
  col_index = sample(dim(Y)[2],20)
  Y_miss = Y[row_index,col_index]

  Y_input = Y
  Y_input[row_index,col_index] = NA
  Y_bs_input = Y_bs
  Y_bs_input[row_index,col_index] = NA

  g1 = greedy(Y_input, K  = 60)
  g2 = greedy(Y_bs_input, K  = 60)

  Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
  results[i,1] =  sqrt(mean((Y_miss - Y_hat_1)^2))
  Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
  results[i,2] = sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
}
saveRDS(results, "./results.rds")
```

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```