### 2017-06-20 analysis/Obj_summary.Rmd
Files commited by wflow_commit.

---
title: "simple summary of objective function for rank 1 vs rank 0"
author: "Wei Wang"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## rank 0 vs rank 1

this document is based on Matthew's suggestion:

> If you simulate rank 0, and run greedy with K=1:
>
> - how often does the K=1 converge to a point mass at 0?
>
> - how often does the K=1 converge to a lower bound that is worse than the lower bound for K=0?
>
> Then you could do the same question for method =fdr (but with this penalized objective function).

### shrink method without penalty term in objective function

we find that we might need fdr method and we should use the objective funciton with penalty.[rank 0 vs rank 1](Objcheck_1.html)

### check the objective function with penalty

In this case, we check the objective funciton with penalty term, and it does monotonically increase.

### fdr method with penalty term in objective function

we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger. [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html)

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```