### 2017-06-21 analysis/Obj_summary.Rmd
Files commited by wflow_commit.

---
title: "simple summary of objective function for rank 1 vs rank 0"
author: "Wei Wang"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## rank 0 vs rank 1

this document is based on Matthew's suggestion:

> If you simulate rank 0, and run greedy with K=1:
>
> - how often does the K=1 converge to a point mass at 0?
>
> - how often does the K=1 converge to a lower bound that is worse than the lower bound for K=0?
>
> Then you could do the same question for method =fdr (but with this penalized objective function).

### shrink method without penalty term in objective function

we find that we might need fdr method and we should use the objective funciton with penalty.[rank 0 vs rank 1](Objcheck_1.html)

### check the objective function with penalty

In this case, we check the objective funciton with penalty term, and it does monotonically increase. [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html)

**now we have two methods (fdr, shrink), and both of them have monotonically increasing corresponding objective function in each case.**

> i) fit flash (with or without penalty)
>
> ii) check if flash-obtained objective function exceeds the K=0 objective. (note: not the limit of the objective as g_f, g_l -> 0, but the actual objective at g_f=g_l=0). If not, set result to g_f=g_l = 0. If so, keep g_f, g_l as output by flash.
>
> -- Matthew's advice.

- we  have done the (i), and we found that it is reasonable to do the (ii). This is because we find that $diff(LB(fitted) - LB(rank=0))$ is negatively correlated with RMSE, which means larger obj value goes with smaller RMSE. So it makes sense we cut the obj value off at the rank 0 model.
- (ii) is easy to implement.

#### questions

1. how to deal with $(\lambda_0 - 1) log \pi_0$ when $\pi_0 = 0$?
2. now we only have two cases: shrink with $\lambda_0 = 1$ and fdr with $\lambda_0 = 10$. Do we need more choice of $\lambda_0$?

### fdr method with penalty term in objective function

we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger. [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html)

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```