### 2017-09-16 analysis/index.Rmd
Files commited by wflow_commit.

---
title: "Home"
output:
  html_document:
    toc: false
---

### simulation studies

### Orthogonal CV

Click on this [orthogonal CV](miss_position.html) to see my example.

Click on this [Questions on Algorithm](first-analysis.html) to see more details.

### flash on the chipseq data from Kevin

Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.

[ns](draft.html)

Click on this [load reslut of GTEx data](GTExresult.html) to see more details.

Click on this [pairwise plot](pairwiseplotofflashgtexzscore.html) to see more details.

### Issues

Click on this [gap issue](ObjGap.html) to see more details. (test)

Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.

- Click on this [rank 0 vs rank 1](Objcheck_1.html) to see more details.
- Click on this [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html) to see more details.

## flashr2

This is a simple test [first try](flash_beta_test.html)