### 2017-09-17 analysis/flash_beta_test.Rmd
Files commited by wflow_commit.

---
title: "test flashr2"
author: "Wei"
date: 09-12
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

### this is code for simulated data

```{r}
sim_hd = function(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0){

  E = matrix(rep(0,N*P),nrow=N)
  sig2_true = matrix(rep(0,N*P),nrow=N)
  for(i in 1:N){
    for(j in 1:P){
      sig2_true[i,j] = mu + a[i] * b[j]
      E[i,j] = rnorm(1,0,sqrt(mu + a[i] * b[j]))
    }
  }

  K=1
  lstart = rnorm(N, 0, signal)

  fstart = rnorm(P, 0, signal)

  index = sample(seq(1:N),(N*SL))
  lstart[index] = 0
  index = sample(seq(1:P),(P*SF))
  fstart[index] = 0

  Y = lstart %*% t(fstart) + E

  return(list(Y = Y, L_true = lstart, F_true = fstart, Error = E,sig2_true = sig2_true))
}
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))

  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)

    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0

    L_true[,k] = lstart
    F_true[,k] = fstart

    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}
```

```{r}
library(flashr2)
```

## check the prediction performance, this is something I don't understand

In this case I run flash for different situation and I found the old version has smaller RMSE.

I print the obj value to check if it is increase.

```{r,message=FALSE,results=FALSE}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 1,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

```{r}
data = set_flash_data(Y)
g1= flash_r1(data,verbose=F)
gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

### for this case, when data are missing, the obj is not increase

the final result is rank zero, so I am not sure it is a problem

```{r}
data.miss = set_flash_data(Y_miss)
g1.miss = flash_r1(data.miss,verbose=F)
```

## missing value

In this case, I have two problems:

1. The Obj function seems not increasing.

2. the RMSE seems a bit large for flashr2. I think the $\tau$ estimation is not that precise.

```{r}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 3,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

```{r}
data.miss = set_flash_data(Y_miss)
g1.miss = flash_r1(data.miss,verbose=F)
gg1.miss = flashr::flash(Y_miss,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1.miss$EL %*% t(g1.miss$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

## compare versions

In this case, we set flash use `shrink` method and compare with the new version.

```{r}
sparse_level_l = seq(0.1,0.9,0.2)
sparse_level_f = seq(0.1,0.9,0.2)
signal_level = seq(1,5)
RMSE_1 = rep(NA,5*5*5)
RMSE_2 = rep(NA,5*5*5)
for(i in 1:5){
  for(j in 1:5){
    for(k in 1:5){
      set.seed(999)
      N = 100
      P = 200
      dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
                  SL = sparse_level_l[j], 
                  signal = signal_level[k],
                  noise = 5)
      Y = dat$Y
      data = set_flash_data(Y)
      g1= flash_r1(data,verbose=F)
      gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
      RMSE_1[25*(i-1) + 5*(j-1) + k] = 
        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
      RMSE_2[25*(i-1) + 5*(j-1) + k] = 
        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
    }
  }
}

plot(RMSE_1 - RMSE_2)
```

To check the extreme case, and I found it is because flashr doesn't output rand zero estimation where zero estimation is better.

```{r}
i = 5
j = 5
k = 1
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
            SL = sparse_level_l[j], 
            signal = signal_level[k],
            noise = 5)
Y = dat$Y
data = set_flash_data(Y)
g1= flash_r1(data,verbose=F)
gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))

  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))

  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))

```
## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```