### 2017-09-29 analysis/applyflahr2onsimulations.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 on the paper"
author: "Wei Wang"
date: 09-28
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## rank 1 data set

### simulated data

```{r,echo=FALSE}
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true))
}
# I think we can try some situation
# pve is small due to the

# this can be a example of the bi sparse factor and loadings

CVPMD=function(Y,c_u,c_v){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_u = length(c_u)
  n_v = length(c_v)
  CVRMSE = array(0,dim = c(n_u,n_v))
  minrmse = Inf
  opt_u = 0
  opt_v = 0
  for(t_u in 1:n_u){
    for(t_v in 1:n_v){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v])
        misshat = out$d * out$u %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_u,t_v] = mean(rmse)
      if(CVRMSE[t_u,t_v] < minrmse){
        minrmse = CVRMSE[t_u,t_v]
        opt_u = c_u[t_u]
        opt_v = c_v[t_v]
      }
    }
  }
  return(list(opt_u = opt_u,opt_v = opt_v))
}

PMA.wrapper = function(Y,ngrids = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_u = seq(1,sqrt(N),length.out = ngrids)
  c_v = seq(1,sqrt(P),length.out = ngrids)
  cvout = CVPMD(Y,c_u,c_v)
  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v )
  residual_PMD = Y - out$d * out$u %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}
```

### sparse

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
#Data = datamaker(N,P,L_pi,L_se,0.36,c(1),c(1),1,sqrt(1))
dim(Data$Y)
# = flashr::flash(Data$Y)
data = flashr2::flash_set_data(Data$Y)
gf = flashr2::flash_r1(data,verbose=FALSE)

PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
MSE_PMD = sqrt(mean(( gPMD$d *(gPMD$u %*% t(gPMD$v)) - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

library(ggplot2)
library(grid)
# zoom bounding box
xlim <- c(-0.08,0.08); ylim <- c(-0.08,0.08)

mat = cbind(Data$L_true,gf$EL,gPMD$u)
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD")
df1 = data.frame(mat1)

base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

g1 <- ggplotGrob(zoomed1)
base1 +annotation_custom(grob = g1, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)

base2 <- ggplot(df1, aes(x = truth, y = PMD)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed2 = ggplot(df1, aes(x = truth, y = PMD)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey() + geom_abline(intercept = 0,slope = 1)

g2 <- ggplotGrob(zoomed2)
base2 +annotation_custom(grob = g2, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)
plot_grid(base1, base2, zoomed1,zoomed2, labels = "AUTO", ncol = 2)

gsfamix_1 = readRDS("../data/simulation/rankone/highsparse/sfamix_res.rds")
gsvd = svd(Data$Y)
gsfa = readRDS("../data/simulation/rankone/highsparse/sfa_res.rds")

# to get the KSVD result
# try other methods
library("R.matlab")
# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
# write the data into matlab
# writeMat("../data/simulation/rankone/highsparse/test_1.mat", data = Data)
# KSVD
res_KSVD = readMat("../data/simulation/rankone/highsparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
Dic = res_KSVD$result[[1]]
loading_Dic = res_KSVD$result[[2]]
MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

# add ssvd into comparison 
gssvd = ssvd::ssvd(Data$Y,method = "method")
MSE_SSVD = sqrt(mean(( gssvd$d * (gssvd$u %*% t(gssvd$v)) - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

library(ggplot2)
library(grid)

mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(Dic),gssvd$u)
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","KSVD","SSVD")
df1 = data.frame(mat1)
xybound = c( -0.75, -0.05, -0.05, 0.55)
z_size = c(-0.08,0.08,-0.08,0.08)
zoom_plot = function(df1,i,zoom_size,bound){
  xlim <- c(zoom_size[1],zoom_size[2])
  ylim <- c(zoom_size[3],zoom_size[4])

  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
    geom_point()  + labs(y = colnames(df1)[i]) +
    theme_bw()+ geom_abline(intercept = 0,slope = 1) + coord_cartesian(xlim=c(-0.65,0.5), ylim=c(-0.65,0.5))
  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

  g1 <- ggplotGrob(zoomed1)
  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
  return(pic1)
}

pic_list = list() 
for(i in 2:7){
  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
}
plot_grid(pic_list[[1]], pic_list[[2]], 
          pic_list[[3]], pic_list[[4]], 
          pic_list[[5]], pic_list[[6]],  labels = "AUTO", ncol = 3)
```

### intermediate sparse

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
dim(Data$Y)
data = flashr2::flash_set_data(Data$Y)
gf = flashr2::flash_r1(data,verbose=FALSE)
PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })

gsfamix_1 = readRDS("../data/simulation/rankone/mediansparse/sfamix_2.rds")
gsfamix_2 = readRDS("../data/simulation/rankone/mediansparse/sfamix_3.rds")
gsfa = readRDS("../data/simulation/rankone/mediansparse/sfa_res.rds")

# to get the KSVD result
# try other methods
library("R.matlab")
# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
# write the data into matlab
# writeMat("../data/simulation/rankone/mediansparse/test_1.mat", data = Data)
# KSVD
res_KSVD = readMat("../data/simulation/rankone/mediansparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
Dic = res_KSVD$result[[1]]
loading_Dic = res_KSVD$result[[2]]
MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

# add ssvd into comparison 
gssvd = ssvd::ssvd(Data$Y,method = "method")

library(ggplot2)
library(grid)
# zoom bounding box
xlim <- c(-0.1,0.1); ylim <- c(-0.1,0.1)

mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","SSVD","KSVD")
df1 = data.frame(mat1)

base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

g1 <- ggplotGrob(zoomed1)
pic1 = base1 +annotation_custom(grob = g1,  xmin = -0.35, xmax = 0, ymin = 0.05, ymax = 0.4)

# mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gssvd$u),as.vector(Dic))
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","SSVD","KSVD")
df1 = data.frame(mat1)
xybound = c(-0.42,0,0.03,0.38)
z_size = c(-0.07,0.07,-0.07,0.07)
zoom_plot = function(df1,i,zoom_size,bound){
  xlim <- c(zoom_size[1],zoom_size[2])
  ylim <- c(zoom_size[3],zoom_size[4])

  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
    geom_point()  + labs(y = colnames(df1)[i]) +
    theme_bw()+ geom_abline(intercept = 0,slope = 1) + coord_cartesian(xlim=c(-0.36,0.36), ylim=c(-0.36,0.36))
  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

  g1 <- ggplotGrob(zoomed1)
  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
  return(pic1)
}

pic_list = list() 
for(i in 2:7){
  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
}
plot_grid(pic_list[[1]], pic_list[[2]], 
          pic_list[[3]], pic_list[[4]], 
          pic_list[[6]], pic_list[[5]],  labels = "AUTO", ncol = 3)
```

## rank 3 case

```{r}

CVPMD=function(Y,c_s,K){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_s = length(c_s)
  CVRMSE = rep(NA,n_s)
  minrmse = Inf
  opt_s = 0
  # here we only need one layer
  for(t_s in 1:n_s){
    rmse = rep(0,5)
    for(i in 1:5){
      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
      misshat = out$u %*%  diag(out$d) %*% t(out$v)
      for(j in 1:5){
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
      }
      rmse[i] = sqrt(rmse[i] / (N * P/5))
    }
    CVRMSE[t_s] = mean(rmse)
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }

  return(list(opt_s = opt_s ))
}

PMA.wrapper = function(Y,ngrids = 10,K){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.8,len=ngrids)
  cvout = CVPMD(Y,c_s, K = K)
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}

```

```{r}
set.seed(99)
N = 150
P = 240
K = 3
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[1:80,1] = rnorm(80,0,0.5)
F_true[81:160,2] = rnorm(80,0,1)
F_true[161:240,3] = rnorm(80,0,2)
L_true[1:10,1] = rnorm(10,0,2)
L_true[11:60,2] = rnorm(50,0,1)
L_true[61:150,3] = rnorm(90,0,0.5)
G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,2),nrow=N)
Y = L_true %*% t(F_true) + E

data = flashr2::flash_set_data(Y)
f_greedy = flashr2::flash_add_greedy(data,Kmax=10)
g_b_flash = flashr2::flash_backfit(data,f_greedy)

outcv = PMA.wrapper(Y,10,K = 3)
library(ssvd)
gssvd = ssvd(Y,r = 3)
```

```{r}
library(reshape2)
melted_cormat_1 <- melt(G, na.rm = TRUE)
melted_cormat_2 <- melt(g_b_flash$EL %*% t(g_b_flash$EF), na.rm = TRUE)
melted_cormat_3 <- melt(outcv$u %*% diag(outcv$d) %*% t(outcv$v), na.rm = TRUE)
melted_cormat_4 <- melt(Y, na.rm = TRUE)
melted_cormat_5 <- melt(gssvd$u %*% diag(gssvd$d) %*% t(gssvd$v), na.rm = TRUE)
# Heatmap
library(ggplot2)
p1 = ggplot(data = melted_cormat_1, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p2 = ggplot(data = melted_cormat_2, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "FLASH",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p3 = ggplot(data = melted_cormat_3, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "PMD",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p4 = ggplot(data = melted_cormat_4, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-5,5), space = "Lab") + labs(title = "observed data",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p5 = ggplot(data = melted_cormat_5, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-5,5), space = "Lab") + labs(title = "SSVD",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

library(cowplot)
plot_grid(p1,p5,p3,p2, ncol = 2)

```

we notice the the flashr2 provide 4 factors in this case while flashr provides 3 factors. I think it is because we use FDR method in flashr method.

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```