### 2017-09-29 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

---
title: "Template for GTEx example on cluster"
author: "wei"
date: 09-29
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Install the package

1 please git clone the `flashr2` on PPS.

`git clone https://github.com/stephenslab/flashr2.git`

2 install `flashr2` locally

`R CMD build flashr2`

`R CMD INSTALL flashr2_0.2-2.tar.gz`

## prepare the data

### original data

Please read your original data as a $N \times P$ matrix in R.

In this case you can create a folder the restore the data as `.rds` file

```
mkdir testflashr2
cd testflashr2
cp gtexEQTL_zscore.rds GTEX/testflashr2/
```

now you have the data matrix in `gtexEQTL_zscore.rds`

to restore the result and track the errors, you can crest folders

```
mkdir output
mkdir ourlog
```

## run FLASH

Here I take the GTEx EQTL zscore as an example

### creat a .R file

```
library(ashr)
library(flashr2)
load("./gtexEQTL_zscore.rds")
Y = t(zscore)
data = flash_set_data(Y)
f_greedy = flash_add_greedy(data,Kmax=60)
f_greedy_bf = flash_backfit(data,f_greedy)
#f_greedy = flash_add_greedy(data,Kmax=60,var_type = "by_column",ash_param=list(method = "fdr"))
#f_greedy_bf = flash_backfit(data,f_greedy,var_type = "by_column",ash_param=list(method = "fdr"))
saveRDS(f_greedy,file = "./output/gflashvarcol.rds")
saveRDS(f_greedy_bf,file = "./output/bflashvarcol.rds")
```

here 60 is much larger than the sample size. our method doesn't restrict $K < \min(P,N)$. But for this case, 60 is enough. We call this file as `flashwrapper.R`

### creat a .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.err
#SBATCH --array=1
#SBATCH --time=32:00:00
#SBATCH --partition=mstephens
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --mem-per-cpu=20G

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/GTEX/testflashr2
Rscript --verbose flashwrapper.R
```

you can substitute the folder path `/home/weidong/HG/flash/data/GTEX/testflashr2`. and we call this `.sbatch` file as `runflash.sbatch`

## run

```
sbatch runflash.sbatch 
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```