### 2017-10-06 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

---
title: "Template for GTEx example on cluster"
author: "wei"
date: 09-29
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Install the package

1 please git clone the `flashr2` on PPS.

`git clone https://github.com/stephenslab/flashr2.git`

2 install `flashr2` locally

`R CMD build flashr2`

`R CMD INSTALL flashr2_0.2-2.tar.gz`

## prepare the data

### original data

Please read your original data as a $N \times P$ matrix in R.

In this case you can create a folder the restore the data as `.rds` file

```
mkdir testflashr2
cd testflashr2
cp gtexEQTL_zscore.rds GTEX/testflashr2/
```

now you have the data matrix in `gtexEQTL_zscore.rds`

to restore the result and track the errors, you can crest folders

```
mkdir output
mkdir ourlog
```

## run FLASH

Here I take the GTEx EQTL zscore as an example

### creat a .R file

```
library(ashr)
library(flashr2)
load("./gtexEQTL_zscore.rds")
Y = t(zscore)
data = flash_set_data(Y)
f_greedy = flash_add_greedy(data,Kmax=60)
f_greedy_bf = flash_backfit(data,f_greedy)
#f_greedy = flash_add_greedy(data,Kmax=60,var_type = "by_column",ash_param=list(method = "fdr"))
#f_greedy_bf = flash_backfit(data,f_greedy,var_type = "by_column",ash_param=list(method = "fdr"))
saveRDS(f_greedy,file = "./output/gflashvarcol.rds")
saveRDS(f_greedy_bf,file = "./output/bflashvarcol.rds")
```

here 60 is much larger than the sample size. our method doesn't restrict $K < \min(P,N)$. But for this case, 60 is enough. We call this file as `flashwrapper.R`

### creat a .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.err
#SBATCH --array=1
#SBATCH --time=30:00:00
#SBATCH --partition=mstephens
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --mem-per-cpu=8G

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/GTEX/testflashr2
Rscript --verbose flashwrapper.R
```

you can substitute the folder path `/home/weidong/HG/flash/data/GTEX/testflashr2`. and we call this `.sbatch` file as `runflash.sbatch`

## run

```
sbatch runflash.sbatch 
```

## plot

```{}
b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
load("../data/GTExdata/gtexEQTL_zscore.rds")
ssY = sum(zscore^2)
K = dim(b_flash$EL)[2] -1
pve = (sapply(seq(1,K),function(x){ sum(b_flash$EL[,x]^2 %*% t(b_flash$EF[,x]^2)) }))/ssY
pve = pmax(round(pve,3),0.001)

dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]

L = b_flash$EL[,1:14]
library(reshape2)
data_L = melt(L)
colnames(data_L) = c("tissue","loading","value")
library(ggplot2)
tissue_color = as.character(colordata[,2])
data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))

ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
  geom_bar(stat = "identity",width = 0.6) +
  scale_fill_manual(values=tissue_color) +
  scale_x_discrete(labels = NULL) +
  theme_grey()+
  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
  facet_wrap(~loading, ncol = 2, scales = "free_y") +
  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
  ggsave("flashr2GTEx1.pdf", width = 8, height = 11)

# the 27th factor is zero
L = b_flash$EL[,15:26]
library(reshape2)
data_L = melt(L)
colnames(data_L) = c("tissue","loading","value")
library(ggplot2)
tissue_color = as.character(colordata[,2])
data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
data_L$loading = factor(data_L$loading,levels = 1:12 ,labels = paste("loading",15:26,"; pve:", pve[15:26]))

ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
  geom_bar(stat = "identity",width = 0.6) +
  scale_fill_manual(values=tissue_color) +
  scale_x_discrete(labels = NULL) +
  theme_grey()+
  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
  facet_wrap(~loading, ncol = 2, scales = "free_y") +
  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
  ggsave("flashr2GTEx2.pdf", width = 8, height = 10)
```

the plots are

<img src="figure/flashr2GTEx1.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">

<img src="figure/flashr2GTEx2.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">

## an example to run SFA on RCC or PPS

1. go to the folder where the data is.

### sfa wrapper

this example is on PPS cluster

```{r}
SFA.wrapper = function(X, K){
  N = dim(X)[1]
  P = dim(X)[2]

  if(file.exists("SFAout")){
    unlink("SFAout", recursive= T)
  }
  system("mkdir SFAout")

  write.table(X,file="dscsfa.txt",row.names=F,col.names=F)
  # sfa command line on cluster
  # ~/flash/simulation/methods/sfa/src/sfa -gen ./dscsfamix.txt -g 600 -k 1 -n 200 -iter 100 -rand 999 -o ./SFAout/sfa

  # sqrt(mean(( loadings %*% factor - test$L_true%*%t(test$F_true))^2))/sqrt(mean((test$L_true%*%t(test$F_true))^2))
  system(paste("~/flash/simulation/methods/sfa/src/sfa", "-gen", "./dscsfa.txt", "-g", N, "-k", K,
               "-n", P, "-iter", 100, "-rand", 999, "-o", "./SFAout/sfa", sep=" "))

  if(file.info("./SFAout/sfa_F.out")$size == 1){
    return(list(L = 0, F = 0))
  } else{
    Fhat=read.table("./SFAout/sfa_F.out")
    lambda=read.table("./SFAout/sfa_lambda.out")
    Fhat=as.matrix(Fhat)
    lambda=as.matrix(lambda)
    return(list(L = lambda, F = Fhat))
  }

}

SFAmix.wrapper = function(X,K){
  N = dim(X)[1]
  P = dim(X)[2]

  write.table(X,file="dscsfamix.txt",row.names=F,col.names=F)
  print(class(X))
  print(dim(X))

  if(file.exists("SFAmixout")){
    unlink("SFAmixout", recursive= T)
  }
  system("mkdir SFAmixout")

  # ~/mvash/mvsim/SFAmix/SFAmix --nf K --y dscrsfamix.txt --out SFAmixout --sep space
  # this is run on the PPS cluster
  system(paste("~/mvash/mvsim/SFAmix/SFAmix","--nf",K,"--y","dscsfamix.txt","--out",
               "SFAmixout","--sep","space",sep=" "))

  alpha=read.table("./SFAmixout/PSI")

  if(file.info("./SFAmixout/EX")$size == 1){
    return(list(L = 0, F = 0))
  } else{
    Fhat=read.table("./SFAmixout/EX")
    lambda=read.table("./SFAmixout/LAM")
    Psi=as.vector(alpha)
    Fhat=as.matrix(Fhat)
    lambda=as.matrix(lambda)
    P=dim(lambda)[2]
    n=dim(Fhat)[1]
    return(list(L = Fhat, F = lambda))
  }

}

```

this is `sfawrapper.R`

### creat run_sfa.R file

```
load("~/flash/dataanalysis/GTeX/EQTLzscore/gtexEQTL_zscore.rds")
setwd("~/flash/dataanalysis/GTeX/EQTLzscore/")
source("./sfawrapper.R")
source("./SFAmixwrapper.R")
Y = zscore
N = dim(Y)[1]
P = dim(Y)[2]
gsfa = SFA.wrapper(Y, 30)
saveRDS(gsfa, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfaGTExzscore.rds")

Y = t(zscore)
gsfamix = SFAmix.wrapper(Y,30)
saveRDS(gsfamix, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfamixGTExzscore.rds")

```

and run it

```
R CMD BATCH run_sfa.R
```

#### remark

For `Y = t(zscore)`, SFA provide NaN result when set `K = 30` and `K = 26`. (all the $L_i$ and $F_j$ are NaN). So I use `Y = zscore` which works in `K = 30` and `K = 26`. In this case, I set `K = 30`

in this case we can get

```
> gsfa = SFA.wrapper(Y, 30)
 Setting G to be 16069
 Setting N to be 44
Trying to read in matrix with G=16069 and N =44
Opening matrix ./dscsfa.txt with 16069 rows and 44 columns
Opening matrix ./dscsfa.txt with 16069 rows and 44 columns
read in matrix 16069 by 44
initializing sfa...
G = 16069
N = 44
K = 30

```

But for `SFAmix` we should use `Y = t(zscore)`.

The approaches make sure we can get $(\hat{\sigma}_1^2,\cdots,\hat{\sigma}_j^2,\cdots)$ is a `16069` vector.

### SFA K=3

```{r}
g3 = readRDS('../data/GTExdata/SFAres/sfaK_3.rds')
par(mfrow = c(1,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
for(k in 1:3){
  barplot(g3$F[k,],main = paste("factor",k))
}
```

### plot SFA result K = 30

```{r}
sfares = readRDS("../data/GTExdata/SFAres/sfaGTExzscore.rds")
# the F is K by N matrix! not N by K matrix
par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
for(k in 1:30){
  barplot(sfares$F[k,],main = paste("factor",k))
}

```

### plot of SFA result with t(Y) K = 10

If you would like to use `t(Y)` you should use `-vn` in the `sfa` command line in order to get $(\hat{\sigma}_1^2,\cdots,\hat{\sigma}_j^2,\cdots)$ is a `16069` vector

```{r}
gt10 = readRDS('../data/GTExdata/SFAres/sfaTYK_10.rds')

par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
for(k in 1:10){
  barplot(gt10$L[,k],main = paste("factor",k))
}
```

#### problem in K = 20 and K = 30 in this case.

```
gsl: lu.c:147: ERROR: matrix is singular
Default GSL error handler invoked.
```

### plot SFAmix result

```{r}
sfamixres = readRDS("../data/GTExdata/SFAres/sfamixGTExzscore.rds")
dim(sfamixres$L)
dim(sfamixres$F)
```

this is only a rank one matrix. I tried two time. two runs give rank one matrix.

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```