### 2017-12-04 analysis/compareOptionsOCV.Rmd
Files commited by wflow_commit.

---
title: "compare the options for missing value MSE"
author: "Wei"
date: 2017-12-04
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Movie Lens data

### FDR without null check vs default setting

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(c("flash_fdr","flash","NBSFA_fdr","NBSFA","PMD_fdr","PMD","softImpute_fdr","softImpute"), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}
data1 = readRDS("../data/cmpOptions/ML100K/FDRnocheck_box.rds")
data2 = readRDS("../data/cmpOptions/ML100K/box_defaultseeting.rds")
data1 = matrix(as.numeric(data1),ncol = 4)
data2 = matrix(as.numeric(data2),ncol = 4)
pM = plot_res(cbind(data1[,1],data2[,1],
                    data1[,2],data2[,2],
                    data1[,3],data2[,3],
                    data1[,4],data2[,4]),"Movie Lens data different settings")
pM
```

We run the experiments twice. `flash_fdr` `NSFA_fdr` `PMD_fdr` and `softImpute_fdr` stand for the results from the simulation data set with FDR method and without Nullcheck in FLASH method. And `flash` `NSFA` `PMD` and `softImpute` represent the results from simulation data set with default setting (shrink, normal with Nullcheck) in FLASH.

## GTEx EQTL Z-score

### FDR without null check vs default setting

```{r}
data1 = readRDS("../data/cmpOptions/GETxZscore/FDRnocheck_box.rds")
data2 = readRDS("../data/cmpOptions/GETxZscore/DefaultSetting_box.rds")
data1 = matrix(as.numeric(data1),ncol = 4)
data2 = matrix(as.numeric(data2),ncol = 4)
pG = plot_res(cbind(data1[,1],data2[,1],
                    data1[,2],data2[,2],
                    data1[,3],data2[,3],
                    data1[,4],data2[,4]),"GTEx z-score different settings")
pG
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```