### 2017-12-04 analysis/GSEA_flashr.Rmd

---
title: "GSEA for flash factors"
author: "Wei"
date: 10-2017
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## try the 3rd factor

Here, we should preare the list of the gene names.

```{r}
res = readRDS("../data/GTEX/bflashvarcol.rds")
load("../data/GTEX/gtexEQTL_zscore.rds")
sorted_f3 = sort(res$EF[,3],index.return = T)
negative_index = sorted_f3$ix[which(sorted_f3$x<0)]
negative_name = row.names(zscore)[negative_index]
new_names = sapply(negative_name, function(x){(strsplit(x,"\\.")[[1]])[1]})
# write.table(new_names,file = "../data/GTEX/f_3_neg.txt",row.names = F,col.names = F)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```