### 2017-12-09 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

---
title: "MSE for rank_1 simulation"
author: "wei wang"
date: 17-12-09
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{}
flash.wrapper = function(Y_data){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_r1(data,verbose=F)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

#here I use SVD instead of KSVD.
SVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gsvd = svd(Y)
  Y_hat = gsvd$d[1] * (gsvd$u[,1] %*% t(gsvd$v[,1]))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SSVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gssvd = ssvd::ssvd(Y,method = "method")
  Y_hat = gssvd$d * (gssvd$u %*% t(gssvd$v))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}
```

```{}
library(PMA)
library(flashr2)
library(ssvd)
source("~/HG/flash/data/simulations/rankone/MSE/Rfunction.R")
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
# Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
RMSE = rep(NA,6)
RMSE[1] = PMA.wrapper(Y_data)
RMSE[2] = flash.wrapper(Y_data)
RMSE[3] = SFA.wrapper(Y_data, K = 1)
RMSE[4] = SFAmix.wrapper(Y_data,K = 1)
RMSE[5] = SVD.wrapper(Y_data)
RMSE[6] = SSVD.wrapper(Y_data)
saveRDS(result, "./output.rds")
```

$$RMSE =\frac{ \sqrt{(Y-\hat{Y} - E )^2}}{\sqrt{(Y - E )^2}}$$

## batch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=02:00:00
#SBATCH --partition=mstephens
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/simulations/rankone/MSE/median
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/simulations/rankone/MSE/median/run.R
```

## run

```
sbatch Jobs.sbatch
```

## plot the result

```
T = 100
results = matrix(NA,ncol = 6, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  # out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_file)
  results[i,] = try(readRDS(file_name))
}
```

## RMSE sparse case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_sparse.rds")
colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
boxplot(sparse_res,main = "sparse case")
```

## RMSE intermediate sparse case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_itermediate.rds")
colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
boxplot(sparse_res,main = "intermediate sparse case")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```