### 2018-01-03 analysis/compare_ash_ebmn.Rmd
Files commited by wflow_commit.

---
title: "compare the different versions of ebnm methods"
author: "Wei"
date: 2018-01-02
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## for the simulated data

```{r,echo=FALSE}
library("ebnm")
library("ashr")
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true,E = E))
}

```

### sparse loding (simulated data)

we can see there is tiny difference between `flash_r1` and `flash_add_greedy`.
The difference between the `ebnm_pn` and `ebnm_ash` are tiny based on the RMSE ratio.

```{r}
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
library(flashr2)
data = flash_set_data(Data$Y)
f_greedy_pn = flash_add_greedy(data,Kmax=10,ebnm_fn = ebnm_pn)
f_greedy_ash = flash_add_greedy(data,Kmax=10,ebnm_fn = ebnm_ash)
# here flashr2 provide 2 factors but 2nd factor is all zero.
f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
par(mfrow = c(2, 2)) 
plot(f_pn$EL,f_greedy_pn$EL[,1],main = "r1 vs greedy (pn)")
plot(f_pn$EL-f_greedy_pn$EL[,1],main = "difference r1 vs greedy")
# there is tiny difference between greedy algorithm and rank one fucntion
plot(f_ash$EL,f_greedy_ash$EL[,1],main = "r1 vs greedy (ash)")
plot(f_ash$EL+f_greedy_ash$EL[,1],main = "difference r1 vs greedy")
# compare the ash and pn methods
par(mfrow = c(1, 1)) 
plot(f_ash$EL,f_pn$EL,main = "pn vs ash (EL)")
Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
c(RMSE_pn,RMSE_ash)
```

Here we can see that the `ebnm_pn` is slightly better than `ebnm_ash`.

We try repeat this experiment 100 times

```{r}
Sim_RMSE = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  Data = datamaker(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae)
  data = flash_set_data(Data$Y)
  f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
  f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
  Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
  RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
  Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
  RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
  return(c(RMSE_pn,RMSE_ash))
}
```

```{r}
T = 100
RMSE = array(NA,dim = c(T,2))
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
for(i in 1:T){
  RMSE[i,] = Sim_RMSE(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
}
colnames(RMSE) = c("pn","ash")
par(mfrow = c(1, 2)) 
boxplot(RMSE,main = "pn vs ash")
boxplot(RMSE[,1] - RMSE[,2],main = "difference")
```

### intermediate sparse loding (simulated data)

The difference between the `ebnm_pn` and `ebnm_ash` are also tiny based on the RMSE ratio.

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
data = flash_set_data(Data$Y)
f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
c(RMSE_pn,RMSE_ash)
```

Here we can see that the `ebnm_pn` is slightly worse than `ebnm_ash`.

```{r}
T = 100
RMSE = array(NA,dim = c(T,2))
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
for(i in 1:T){
  RMSE[i,] = Sim_RMSE(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
}
colnames(RMSE) = c("pn","ash")
par(mfrow = c(1, 2)) 
boxplot(RMSE,main = "pn vs ash")
boxplot(RMSE[,1] - RMSE[,2],main = "difference")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```