### 2018-01-04 analysis/compare_ash_pn_miss.Rmd
Files commited by wflow_commit.

---
title: "compare the different versions of ebnm methods on missing data"
author: "Wei"
date: 2018-01-02
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
OCV_index=function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold

  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
}
# OCVindex = OCV_index(Y,k_fold = 5)

OCV_data = function(Y,OCVindex,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing= array(0,dim = c(k_fold,N,P))
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  return(missing)
}
# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)

OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing = OCVdata
  SSE = rep(0,k_fold)
  for(i in 1:k_fold){
    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
    for(j in 1:k_fold){
      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
    }
  }
  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
  return(RMSE)
}

call_method = function(Y_data,method,Kmax = 50){
  if(method == "ebnm_pn"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn,verbose = F)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "ebnm_ash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_ash,verbose = F)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else{
    stop("the method is not implemented yet, please check it out")
  }
  return(Y_hat)
}
```

## denoise tumor data example in our paper

```{r}
set.seed(99)
library(denoiseR)
data(tumors)

Y = tumors[, -ncol(tumors)]
Y = as.matrix(Y)
N = dim(Y)[1]
P = dim(Y)[2]

OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 20)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 20)

```

Check the RMSE for the missing value

```{r}
c(f_ash,f_pn)
```

## Breast cancer data example in our paper

```{r}
set.seed(99)
library(R.matlab)
## run the code
Y_centered = readMat("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/Breastcancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 50)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 50)
```

Check the RMSE for the missing value

```{r}
c(f_ash,f_pn)
```

### MovieLens data

```{}
set.seed(11)
library(Matrix)
ml100K_data = readRDS("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/ML100K/biscale_data.rds")
MLMatrix <- sparseMatrix(i = ml100K_data[,1],
                         j = ml100K_data[,2],
                         x = ml100K_data[,3],dims = c(943,1682))
# turn this sparse matrix into matrix in r
Y = as.matrix(MLMatrix)
Y[which(Y == 0)] = NA
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_pn",Kmax = 10)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_ash",Kmax = 10)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```