### 2018-01-10 analysis/PCAresult.Rmd
Files commited by wflow_commit.

---
title: "PCA(SVD) result"
author: "Wei"
date: 2018-01-10
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
plot_GTEXres = function(L,pve){
  library(reshape2)
  data_L = melt(L)
  colnames(data_L) = c("tissue","loading","value")
  library(ggplot2)
  tissue_color = as.character(colordata[,2])
  data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
  data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))

  ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
    geom_bar(stat = "identity",width = 0.6) +
    scale_fill_manual(values=tissue_color) +
    scale_x_discrete(labels = NULL) +
    theme_grey()+
    theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
    labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
    facet_wrap(~loading, ncol = 2, scales = "free_y") +
    guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
}
b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
load("../data/GTExdata/gtexEQTL_zscore.rds")
ssY = sum(zscore^2)
K = dim(b_flash$EL)[2] -1
pve = (sapply(seq(1,K),function(x){ sum(b_flash$EL[,x]^2 %*% t(b_flash$EF[,x]^2)) }))/ssY
pve = pmax(round(pve,3),0.001)

dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
```

```{r}
L = b_flash$EL[,1:8]
plot_GTEXres(L,pve)
L = b_flash$EL[,9:16]
plot_GTEXres(L,pve)

gsvd = svd(zscore)
pve = (sapply(seq(1,K),function(x){ sum((gsvd$d[x] * gsvd$u[,x] %*% t(gsvd$v[,x]))^2) }))/ssY
pve = pmax(round(pve,3),0.001)
L = gsvd$v[,1:8]
plot_GTEXres(L,pve)
L = gsvd$v[,9:16]
plot_GTEXres(L,pve)

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```