### 2018-01-10 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 with OCV"
author: "wei wang"
date: 09-30
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## preparation

1. creat a folder

```
mkdir Breastcancer
```

2. put your data into this folder

here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.

3. creat the wrapper function in R and matlab

### wrapper function for OCV

```{r}
OCV_index=function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold

  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
}
# OCVindex = OCV_index(Y,k_fold = 5)

OCV_data = function(Y,OCVindex,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing= array(0,dim = c(k_fold,N,P))
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  return(missing)
}
# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)

OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing = OCVdata
  SSE = rep(0,k_fold)
  for(i in 1:k_fold){
    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
    for(j in 1:k_fold){
      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
    }
  }
  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
  return(RMSE)
}

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ash_param=list(method = "fdr"),var_type = "constant")
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "PMD"){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == "softImpute"){
    gsoft = softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else{
    stop("the method is not implemented yet, please check it out")
  }
  return(Y_hat)
}

NSF_OCV = function(OCVindex){
  if(file.exists("NBSFAout")){
    unlink("NBSFAout", recursive= T)
  }
  system("mkdir NBSFAout")
  # write the missing index

  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)

  system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\'); run(\'missvalue.m\');exit;\" ")
  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
  nsfa_rmse = as.vector(g_nsfa$RMSE)
  return(nsfa_rmse)
}

```

Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`

### matlab code

```{}
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat;
Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

load(fullfile(pwd,'NBSFAout/OCVindex.mat'))

missindex = ones(settings.D,settings.N);
k_fold = 5;
foldindex = zeros(k_fold,k_fold,2);
for k = 1:2
    for j = 1:k_fold
        for i = 1:k_fold
            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
        end
    end
end

SSE = zeros(1,k_fold);
for i = 1:k_fold
    mvmask = missindex;
    for j = 1:k_fold
        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
    end
    initialsample=init_nsfa(settings);
    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
    Y_hat = (finalsample.G *  finalsample.X);
    for j = 1:k_fold
        SSE(i) = SSE(i) + ...
            sum(sum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
    end

end

RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
RMSE;

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'RMSE'); 
```

you need change the path as well. this file name is 'missvalue.m'

### creat the run.R file to run the code

In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.

```{}
library(R.matlab)
## run the code
Y_centered = readMat("~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "softImpute",Kmax = 50)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "PMD",Kmax = 50)
nsfa_rmse = NSF_OCV(OCVindex)
flashG_rmse = mean(c( nsfa_rmse,pmd_rmse,soft_rmse))

result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")

```

this file name is `runOCV.R`

### creat folder to track the result

```
mkdir outlog
```

### creat .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=06:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/runOCV.R
```

this file is name as `OCV.sbatch`

## run OCV

now in your folder, you should have:

```
[weidong@midway2-login1 Breastcancer]$ ls
OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
```

```
sbatch OCV.sbatch
```

## to get the result

```{}
T = 100
results = matrix(NA,ncol = 22, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./boxplot.rds")
```

## plot the result

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}

PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
pp = plot_res(PT_res,"Text data")
DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
pd = plot_res(DT_res,"Tumor data")
GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
pg = plot_res(GZ_res,"GTEx data")
BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
pb = plot_res(BC_res,"Breast Cancer data")
ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
pM = plot_res(ML_res,"Movie Lens data")
```

## plots

```{r}
gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
```

## code for the tuning parameters from OCV only change one parameter

try this one later.....

```{r}
CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD",N_sim = 10){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    for(iter in 1:N_sim){
      # each time we set the rmse to zeros
      rmse = rep(0,fold)
      for(i in 1:fold){
        if(method == "PMD"){
          out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
          # out = PMD(missing[i,,], K = K)
        }else{
          out = softImpute(Y, rank.max = K,lambda = c_s[t_s]) 
        }
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
        for(j in 1:fold){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
      } #get the result for one run
      print(sqrt(sum(rmse)/(N*P)))
      CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    }
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  CVRMSE = CVRMSE/N_sim
  return(list(opt_s = opt_s, output = CVRMSE))
}

PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.99,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}

softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(softImpute)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,10,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s) 
  residual_softImpute = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_softImpute))
}
```

#### new reuslt

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
  softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
  methods = rep(c("flash","NBSFA",pmd_c,softImpute_c), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}
```

```{r}
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
pt = plot_res(PT_res,"Tumor data")
pt
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
pp = plot_res(PT_res,"Text data")
pp
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```