### 2018-01-25 analysis/setup.Rmd
wflow_publish("setup.Rmd")

---
title: "Setup for flash project"
output:
  html_document:
    toc: false
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

To reproduce our results on your own computer, please follow these
setup instructions.

## Setup instructions

*Fill out setup instructions in this section. See
[here](https://stephenslab.github.io/wflow-divvy/setup.html) and
[here](https://github.com/pcarbo/mouthwash_sims/blob/master/README.md)
for some ideas.*

## Session information

Here are details about the computing environment, including the
version of R and the R packages, that were used to generate the
results from the R Markdown documents.

```{r session-info}
```