### 2018-01-25 analysis/MSErank3sim.Rmd
Files commited by wflow_commit.

---
title: "MSE for rank_1 simulation"
author: "wei wang"
date: 18-01-23
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

This file is just generate the RRMSE plot of rank 3 data.

```{r}
# the example in our paper using sigma = 2
DataMaker = function(sigma = 2){
  N = 150
  P = 240
  K = 3
  L_true = array(0,dim = c(N,K))
  F_true = array(0,dim = c(P,K))
  F_true[1:80,1] = rnorm(80,0,0.5)
  F_true[81:160,2] = rnorm(80,0,1)
  F_true[161:240,3] = rnorm(80,0,2)
  L_true[1:10,1] = rnorm(10,0,2)
  L_true[11:60,2] = rnorm(50,0,1)
  L_true[61:150,3] = rnorm(90,0,0.5)
  G = L_true %*% t(F_true)
  # generate Y
  E = matrix(rnorm(N*P,0,sigma),nrow=N)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y, L_true = L_true, F_true = F_true, E = E))
}

CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    # each time we set the rmse to zeros
    rmse = rep(0,fold)
    for(i in 1:fold){
      if(method == "PMD"){
        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
      }else{
        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
      }
      if(length(out$d)==1){
        misshat = (out$d) * out$u %*% t(out$v)
      }else{
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
      }
      for(j in 1:fold){
        # for each fold j
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
      }
    } #get the result for one run
    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  return(list(opt_s = opt_s, output = CVRMSE))
}

PMA.wrapper = function(Y_data, ngrids = 10, K=3, fold = 10){
  library(PMA)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.1,0.9,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
  Y_hat = out$u %*% diag(out$d) %*%  t(out$v)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

softImpute.wrapper = function(Y_data, ngrids = 10, K = 3, fold = 10){
  library(softImpute)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,100,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
  if(length(out$d)==1){
    Y_hat = (out$d) * out$u %*% t(out$v)
  }else{
    Y_hat = out$u %*%  diag(out$d) %*% t(out$v)
  }
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SVD.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gsvd = svd(Y,nu = 3,nv = 3)
  Y_hat = (gsvd$u[,1:K] %*% diag(gsvd$d[1:K]) %*% t(gsvd$v[,1:K]))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SSVD.wrapper = function(Y_data,K = 3){
  library(ssvd)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gssvd = ssvd::ssvd(Y,method = "method",r = K)
  Y_hat = (gssvd$u %*% diag(gssvd$d) %*% t(gssvd$v))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

PN.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  library(ebnm)
  library(flashr2)
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn, K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

flash.wrapper = function(Y_data, K = 3){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}
```

creat `run.R` file as follows:

```{}
library(PMA)
library(flashr2)
library(ssvd)
library(softImpute)
Y_data = DataMaker()
RMSE = rep(NA,6)
RMSE[1] = PMA.wrapper(Y_data)
RMSE[2] = flash.wrapper(Y_data)
RMSE[3] = PN.wrapper(Y_data)
RMSE[4] = SVD.wrapper(Y_data)
RMSE[5] = SSVD.wrapper(Y_data)
RMSE[6] = softImpute.wrapper(Y_data)
saveRDS(RMSE, "./output.rds")
```

```{}
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=./outlog/arrayJob_%A_%a.out
#SBATCH --error=./outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=02:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

module load R

# Print this sub-job's task ID
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose ../run.R
```

name the above file as `Jobs.sbatch`

```
sbatch Jobs.sbatch
```

## get the result

```
T = 100
results = matrix(NA,ncol = 6, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_file = "output.rds"
  file_name = file.path(test_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./RES_rrmse.rds")
```

## plot the result

```{r}
plot_res = function(output,title = "data",legend_position = "none", methods_name){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(methods_name, each = N)
  df = data.frame(RRMSE_diff = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RRMSE_diff, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=10),
          plot.title = element_text(size = 18, face = "bold"),
          axis.text.y = element_text(size =15), 
          axis.text.x = element_text(size =15,angle = 45, hjust = 1))
  p
}
library(ggplot2)
```

```{r}
sparse_res = readRDS("../data/simulation/rankthree/boxplot.rds")
colnames(sparse_res) = c("PMD.cv1","flash","flash.pn","SVD","SSVD","SI.cv")
sparse_diff = sparse_res - sparse_res[,2]
# p3 = plot_res(sparse_diff,"Difference from FLASH result (0% zeros)")
methods_name = colnames(sparse_diff)
p3 = plot_res(sparse_diff,title = "Rank 3 example",methods_name = methods_name)
```

```{r}
p3
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```