### 2018-01-25 analysis/OCVmissdataSetting.Rmd
Files commited by wflow_commit.

---
title: "Some useful information of OCV for missing data"
author: "wei wang"
date: 2018-01-25
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## settings for the 5 data sets in our paper.

## other experiments during the test

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label){
  rmse = as.vector(output)
  N = dim(output)[1]
  # methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}
x_label = c("flash","NBSFA","PMD","softImpute")
PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
pp = plot_res(PT_res,"Text data",x_label = x_label)
DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
pd = plot_res(DT_res,"Tumor data",x_label = x_label)
GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
pg = plot_res(GZ_res,"GTEx data",x_label = x_label)
BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
pb = plot_res(BC_res,"Breast Cancer data",x_label = x_label)
ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
pM = plot_res(ML_res,"Movie Lens data",x_label = x_label)
```

## plots

```{r}
gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))

```

## results for different tuning parameters

```{r}
pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
x_label= c("flash","NBSFA",pmd_c,softImpute_c)
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
pt = plot_res(PT_res,"Tumor data",x_label = x_label)
pt
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
pp = plot_res(PT_res,"Text data",x_label = x_label)
pp
```

## use OCV choose the tuning parameters

labels | null check | backfitting | greedy | ebnm_ash | ebnm_pn 
------ | ---------- | ----------- | ------ | -------- | ------
flashG | yes| |yes | yes |
flashGwn | | | yes|yes |
flashB | yes| yes| | yes|
PN | yes | | yes| |yes

### Breast Cancer data

we use 10 grids for `softImpute` and `PMD`

```{r}
PB_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
pb = plot_res(PB_res,"Breast Cancer data",x_label = x_label)
pb
```

we take the 'flashB' away to compare the rest.

```{r}
x_label= c("PN","flashG","flashGwn","nsfa","pmd","soft")
pb = plot_res(PB_res[,-4],"Breast Cancer data",x_label = x_label)
pb
```

### denoiseR Tumor data

we use 10 grids for `softImpute` and `PMD`

```{r}
PT_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
pt = plot_res(PT_res,"Tumor data",x_label = x_label)
pt
```

### text data

in this data, we add `zero` as use zero values as imputation to compare with other methods.

```{r}
PT_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
pt = plot_res(PT_res,"Text data",x_label = x_label)
pt
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```