### 2018-01-27 analysis/index.Rmd
Files commited by wflow_commit.

---
title: "Home"
output:
  html_document:
    toc: false
---

## Simulation studies

### Orthogonal CV

Click on this [orthogonal CV](miss_position.html) to see my example.

### Apply flashr2 for the examples in FLASH paper

#### simulations

Click here to check the [simulatoin result (rank one and rank three)](applyflahr2onsimulations.html)

Click here to check the [RMSE for the rank one example in the paper](MSErank1sim.html)

Click here to check the [RMSE for the rank three example in the paper](simulation3.html) and [RRMSE](MSErank3sim.html)

#### GTEx Z-score

Here is a example to run `flashr2` on RCC [RCC template](GTEXzscoreTemplate.html)

compare with PCA result[compare with SVD](PCAresult.html)

#### missing data imputation

click here to check the template for [missing data imputation with OCV](templateforOCV.html) and [CV template](CVtemplate.html)

click here to check the comparison of different options for FLASH [different shrink options](compareOptionsOCV.html)