### 2018-01-27 analysis/oldanalysis.Rmd

---
title: "old experiments"
author: "wei wang"
date: 2018-01-27
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

Click on this [Questions on Algorithm](first-analysis.html) to see more details.

### flash on the chipseq data from Kevin

Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.

[ns](draft.html)

Click on this [load reslut of GTEx data](GTExresult.html) to see more details.

Click on this [pairwise plot](pairwiseplotofflashgtexzscore.html) to see more details.

### Issues

Click on this [gap issue](ObjGap.html) to see more details. (test)

Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.

- Click on this [rank 0 vs rank 1](Objcheck_1.html) to see more details.
- Click on this [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html) to see more details.

## flashr2

This is a simple test [first try](flash_beta_test.html)

Initialization issue: [softimpute as initialization](missingimputeinitial.html)

Click here to find the [null check for flashr2](NULLcheck.html) and [densen nu case](nullcheck2.html)

### different options for EBNM

#### simulation examples in our paper

Click here to check the [simulated examples](compare_ash_ebmn.html)

Click here to check the [missing data examples (first simple comparison)](compare_ash_pn_miss.html)

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```