### 2018-01-27 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

---
title: "MSE for rank_1 simulation"
author: "wei wang"
date: 17-12-09
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## run the methods

### 1. creat a folder

```
mkdir rankone
```

### 2. creat .R file

```{r}
# this file is for all the R function we need here for sparse factor model
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y, L_true = L_true, F_true = F_true, E = E))
}

CVPMD=function(Y,c_u,c_v){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_u = length(c_u)
  n_v = length(c_v)
  CVRMSE = array(0,dim = c(n_u,n_v))
  minrmse = Inf
  opt_u = 0
  opt_v = 0
  for(t_u in 1:n_u){
    for(t_v in 1:n_v){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v],center=FALSE)
        misshat = out$d * out$u %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_u,t_v] = mean(rmse)
      if(CVRMSE[t_u,t_v] < minrmse){
        minrmse = CVRMSE[t_u,t_v]
        opt_u = c_u[t_u]
        opt_v = c_v[t_v]
      }
    }
  }
  return(list(opt_u = opt_u,opt_v = opt_v))
}

# input is a list of Y L_true and F_true which is called Y_data
PMA.wrapper = function(Y_data,ngrids = 10){
  library(PMA)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_u = seq(1,sqrt(N),length.out = ngrids)
  c_v = seq(1,sqrt(P),length.out = ngrids)
  cvout = CVPMD(Y,c_u,c_v)
  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v,center=FALSE )
  Y_hat = out$d * out$u %*% t(out$v)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

PMA_d.wrapper = function(Y_data){
  library(PMA)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  cv.out <- PMA::PMD.cv(Y, type="standard", sumabss=seq(0.1, 1, len=20),center=FALSE)
  out <-  PMA::PMD(Y, type="standard", sumabs=cv.out$bestsumabs, K=1, v=cv.out$v.init,center=FALSE)
  Y_hat = out$d * out$u %*% t(out$v)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

PN.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  library(ebnm)
  library(flashr2)
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

flash.wrapper = function(Y_data){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant")
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SFA.wrapper = function(Y_data, K){
  #sfa put sparse on loadings
  Y = Y_data$Y
  # N vector
  L_true = Y_data$L_true
  # P vector
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]

  if(file.exists("SFAout")){
    unlink("SFAout", recursive= T)
  }
  system("mkdir SFAout")

  write.table(Y, file="dscsfa.txt",row.names=F,col.names=F)
  # sfa command line on cluster
  # ~/flash/simulation/methods/sfa/src/sfa -gen ./dscsfamix.txt -g 600 -k 1 -n 200 -iter 100 -rand 999 -o ./SFAout/sfa

  # sqrt(mean(( loadings %*% factor - test$L_true%*%t(test$F_true))^2))/sqrt(mean((test$L_true%*%t(test$F_true))^2))
  system(paste("~/HG/flash/data/missingvalue/methods/sfa/src/sfa", "-gen", "./dscsfa.txt", "-g", N, "-k", K,
               "-n", P, "-iter", 100, "-rand", 999, "-o", "./SFAout/sfa", sep=" "))

  if(file.info("./SFAout/sfa_F.out")$size == 1){
    L_hat = 0
    F_hat = 0
    RMSE = 1
    return(RMSE)
  } else{
    Fhat=read.table("./SFAout/sfa_F.out")
    lambda=read.table("./SFAout/sfa_lambda.out")
    Fhat=as.matrix(Fhat)
    lambda=as.matrix(lambda)
    L_hat = lambda
    F_hat = Fhat

    Y_hat = L_hat %*% F_hat
    RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
    return(RMSE)
  }
}

SFAmix.wrapper = function(Y_data,K){
  Y = t(Y_data$Y)
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = t(Y_data$E)
  N = dim(Y)[1]
  P = dim(Y)[2]

  write.table(Y,file="dscsfamix.txt",row.names=F,col.names=F)
  print(class(Y))
  print(dim(Y))

  if(file.exists("SFAmixout")){
    unlink("SFAmixout", recursive= T)
  }
  system("mkdir SFAmixout")

  # ~/mvash/mvsim/SFAmix/SFAmix --nf K --y dscrsfamix.txt --out SFAmixout --sep space
  # this is run on the PPS cluster
  system(paste("~/HG/flash/data/missingvalue/methods/SFAmix/SFAmix",
               "--nf",K,"--y","dscsfamix.txt","--out",
               "SFAmixout","--sep","space",sep=" "))

  alpha=read.table("./SFAmixout/PSI")

  if(file.info("./SFAmixout/EX")$size == 1){
    L_hat = 0
    F_hat = 0
    RMSE = 1
    return(RMSE)
  } else{
    Fhat=read.table("./SFAmixout/EX")
    lambda=read.table("./SFAmixout/LAM")
    Psi=as.vector(alpha)
    Fhat=as.matrix(Fhat)
    lambda=as.matrix(lambda)
    # return(list(L = Fhat, F = lambda))
    L_hat = Fhat
    F_hat = lambda

    Y_hat = L_hat %*% F_hat
    RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
    return(RMSE)
  }
}

SVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gsvd = svd(Y)
  Y_hat = gsvd$d[1] * (gsvd$u[,1] %*% t(gsvd$v[,1]))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SSVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gssvd = ssvd::ssvd(Y,method = "method")
  Y_hat = gssvd$d * (gssvd$u %*% t(gssvd$v))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

CV_softImpute=function(Y,c_s,K,fold = 5, method = "PMD"){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    # each time we set the rmse to zeros
    rmse = rep(0,fold)
    for(i in 1:fold){
      if(method == "PMD"){
        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
      }else{
        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
      }
      if(length(out$d)==1){
        misshat = (out$d) * out$u %*% t(out$v)
      }else{
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
      }
      for(j in 1:fold){
        # for each fold j
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
      }
    } #get the result for one run
    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  return(list(opt_s = opt_s, output = CVRMSE))
}

softImpute.wrapper = function(Y_data, ngrids = 10, K = 1, fold = 5){
  library(softImpute)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,100,len=ngrids)
  cvout = CV_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
  Y_hat = out$d * out$u %*% t(out$v)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

```

name this file as `Rfunction.R`

### creat the folder for different case

```{}
mkdir intermediate
```

### creat the run.R file

```{}
cd intermediate
```

creat `run.R` file as follows:

```{}
library(PMA)
library(flashr2)
library(ssvd)
source("../Rfunction.R")
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
# Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))  sparse case
# Data = datamaker(N,P,L_pi,L_se,1.0,c(1),c(1),1,sqrt(25))  dense case
RMSE = rep(NA,6)
RMSE[1] = PMA.wrapper(Y_data)
RMSE[2] = flash.wrapper(Y_data)
RMSE[3] = SFA.wrapper(Y_data, K = 1)
RMSE[4] = SFAmix.wrapper(Y_data,K = 1)
RMSE[5] = SVD.wrapper(Y_data)
RMSE[6] = SSVD.wrapper(Y_data)
saveRDS(result, "./output.rds")
```

The score we restroe is as follows:

$$RMSE =\frac{ \sqrt{(Y-\hat{Y} - E )^2}}{\sqrt{(Y - E )^2}}$$

### batch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=./outlog/arrayJob_%A_%a.out
#SBATCH --error=./outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=02:00:00
#SBATCH --partition=mstephens
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose ../run.R
```

name the above file as `Jobs.sbatch`

```
sbatch Jobs.sbatch
```

## get the result

```
T = 100
results = matrix(NA,ncol = 6, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_file = "output.rds"
  file_name = file.path(test_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./RES_rrmse.rds")
```

## plot the result

## RMSE sparse case 100 simulations

```{r}
plot_res = function(output,title = "data",legend_position = "none", methods_name,myColors){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(methods_name, each = N)
  df = data.frame(RRMSE_diff = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RRMSE_diff, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+ scale_color_manual(values=myColors)+
    theme(legend.position= legend_position, legend.text=element_text(size=10),
          axis.text.y = element_text(size =12),
          axis.text.x = element_text(size =12,angle = 45, hjust = 1))
  p
}
library(ggplot2)
```

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_sparse.rds")
colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
sparse_diff = sparse_res - sparse_res[,2]
methods_name = colnames(sparse_diff)
fill_color = c("red","yellow3", "springgreen", "springgreen3", "cyan","cyan3", "deepskyblue","violet","purple")
p1 = plot_res(sparse_diff,title = "Difference from FLASH result (90% zeros)",
              methods_name = methods_name, myColors = fill_color)
methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
```

## RMSE intermediate sparse case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_itermediate.rds")
colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
sparse_diff = sparse_res - sparse_res[,2]
# p2 = plot_res(sparse_diff,"Difference from FLASH result (30% zeros)")

methods_name = colnames(sparse_diff)
p2 = plot_res(sparse_diff,title = "Difference from FLASH result (30% zeros)",methods_name = methods_name,myColors = fill_color)
methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
```

## RMSE dense case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_dense.rds")
colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
sparse_diff = sparse_res - sparse_res[,2]
# p3 = plot_res(sparse_diff,"Difference from FLASH result (0% zeros)")
methods_name = colnames(sparse_diff)
p3 = plot_res(sparse_diff,title = "Difference from FLASH result (0% zeros)",methods_name = methods_name,myColors = fill_color )
methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
```

```{r}
# gridExtra::grid.arrange(p1,p2,p3,p1_cut,p2_cut,p3_cut,ncol = 3)
gridExtra::grid.arrange(p1,p1_cut,p2,p2_cut,p3,p3_cut,ncol = 2)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```