### 2018-02-11 analysis/RsessionInfo.Rmd
Files commited by wflow_commit.

---
title: "R Session Info"
author: "Wei Wang"
date: 
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
library(flashr)
library(ssvd)
library(methods)
library(PMA)
library(R.matlab)
library(Matrix)
library(softImpute)
library(ggplot2)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```