### 2018-02-11 analysis/gtex_si.rmd
Files commited by wflow_commit.

---
title: "Untitled"
author: "Matthew Stephens"
date: "January 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(1)
Zscores = readRDS("~/Downloads/MatrixEQTLSumStats.Portable.Z.rds")
Ztop = Zscores$test.z
z.s = softImpute::softImpute(Ztop,rank.max=30,lambda=89)
```

```{r}
#this plotting due to Wei Wang
library(ggplot2)
library(reshape2)
#' @param L a ntissue by K matrix of loadings
#' @param pve a vector of K pve values
#' @param offset a number to add to the loading numbers in titles  (eg set to 9 if plotting loadings 9-x)
plot_GTEXres = function(L,pve,offset=0){
  dat = read.table("GTExColors.txt", sep = '\t', comment.char = '')
  nL = dim(L)[2]
  colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
  data_L = melt(L)
  colnames(data_L) = c("tissue","loading","value")
  tissue_color = as.character(colordata[,2])
  data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
  data_L$loading = factor(data_L$loading,levels = 1:nL ,labels = paste("factor",offset+(1:nL),"; pve:", round(pve,3)))

  ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
    geom_bar(stat = "identity",width = 0.6) +
    scale_fill_manual(values=tissue_color) +
    scale_x_discrete(labels = NULL) +
    theme_grey()+
    theme(legend.position="right", legend.text=element_text(size=9), axis.text.y = element_text(size = 5)) + 
    labs(title = "GTEx data", y = "factor values" ,x = "tissues", fill="tissue") +
    facet_wrap(~loading, ncol = 2, scales = "free_y") +
    guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
}

pve.si = z.s$d^2/sum(Ztop^2)
plot_GTEXres(z.s$v[,1:8],pve.si[1:8])
plot_GTEXres(z.s$v[,9:16],pve.si[9:16],9)

```