### 2018-02-11 analysis/setup.Rmd

---
title: "Setup for flash project"
output:
  html_document:
    toc: false
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

To reproduce our results on your own computer, please follow these
setup instructions.

## Setup instructions

### Install software and R packages

1. Install [R](https://cran.r-project.org).

2. Install [MATLAB](https://www.mathworks.com/products/matlab.html)

3. Install the required R packages (run the following commands in the R).

```R
install.packages(c("devtools", "ashr", "PMA","ssvd","softImpute","Matrix","R.matlab","ggplot2","denoiseR"))
devtools::install_github("stephenslab/flashr")
```

4. Install MATLAB package [NBSFA](https://github.com/davidaknowles/nsfa)

5. Install packages [sfa](http://stephenslab.uchicago.edu/software.html#sfa) and [SFAmix](http://beehive.cs.princeton.edu/software/)

### Data sets

All the data sets used in the paper are in [data files](https://github.com/NKweiwang/flash_workflow/tree/master/data/output/missingdata)

## Session information

Here are details about the computing environment, including the
version of R and the R packages, that were used to generate the
results from the R Markdown documents.

```{r,echo=FALSE}
library(flashr)
library(ssvd)
library(methods)
library(PMA)
library(R.matlab)
library(Matrix)
library(softImpute)
library(ggplot2)
library(denoiseR)
```

```{r session-info}
```