### 2018-02-11 analysis/templateforOCV.Rmd

---
title: "apply flashr with OCV"
author: "wei wang"
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## preparation

1. creat a folder

```
mkdir Breastcancer
```

2. put your data into this folder

here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.

3. creat the wrapper function in R and matlab

### wrapper function for OCV

```{r}
CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    # each time we set the rmse to zeros
    rmse = rep(0,fold)
    for(i in 1:fold){
      if(method == "PMD"){
        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
      }else{
        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
      }
      if(length(out$d)==1){
        misshat = (out$d) * out$u %*% t(out$v)
      }else{
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
      }
      for(j in 1:fold){
        # for each fold j
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
      }
    } #get the result for one run
    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  return(list(opt_s = opt_s, output = CVRMSE))
}

PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.3,0.8,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
  return(list(d = out$d, u = out$u, v = out$v))
}

softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
  library(softImpute)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,100,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
  return(list(d = out$d, u = out$u, v = out$v))
}

OCV_index=function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold

  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
}
# OCVindex = OCV_index(Y,k_fold = 5)

OCV_data = function(Y,OCVindex,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing= array(0,dim = c(k_fold,N,P))
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  return(missing)
}
# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)

OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing = OCVdata
  SSE = rep(0,k_fold)
  for(i in 1:k_fold){
    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
    for(j in 1:k_fold){
      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
    }
  }
  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
  return(RMSE)
}

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_wn"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "flash_gb"){
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
    f = flashr::flash_backfit(data,f_greedy)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "flash_bf"){
    data = flashr::flash_set_data(Y_data)
    f_data = flashr::flash_add_factors_from_data(data,K = Kmax)
    f = flashr::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else if(method == "PMD"){
    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
    if(length(out$d)==1){
      Y_hat =  (out$d) * out$u %*% t(out$v)
    }else{
      Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
    }
  }else if(method == 'PN'){
    library(ebnm)
    library(flashr)
    data = flashr::flash_set_data(Y_data)
    f_greedy = flashr::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == 'PMD_d'){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == 'softImpute_d'){
    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else if(method == "softImpute"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    if(length(gsoft$d)==1){
      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
    }else{
      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
    }
  }else if(method == "SF_flash"){
    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
    LL = gsoft$u %*% diag(gsoft$d)
    FF = gsoft$v
    data = flashr::flash_set_data(Y_data)
    f_data = flashr::flash_add_lf(data, LL=LL, FF=FF)
    f = flashr::flash_backfit(data,f_data)
    Y_hat = f$EL %*% t(f$EF)
  }else{
    # stop("the method is not implemented yet, please check it out")
    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
  }
  return(Y_hat)
}

NSF_OCV = function(OCVindex){
  if(file.exists("NBSFAout")){
    unlink("NBSFAout", recursive= T)
  }
  system("mkdir NBSFAout")
  # write the missing index

  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)

  system("matlab -nosplash -nodesktop -r \"addpath(\'../\'); run(\'missvalue.m\');exit;\" ")
  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
  nsfa_rmse = as.vector(g_nsfa$RMSE)
  return(nsfa_rmse)
}

```

This file name is `wrapper.R`

### matlab code

In this folder

```{}
ddpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ../Ydata.mat;
Ycentered=Y;
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

file_path = fullfile(pwd,'NBSFAout/OCVindex.mat');
load(file_path);

% missindex = ones(settings.D,settings.N); this is not missing index this
% is observed index.
missindex = (~isnan(Y));
k_fold = 10;
foldindex = zeros(k_fold,k_fold,2);
for k = 1:2
    for j = 1:k_fold
        for i = 1:k_fold
            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
        end
    end
end

SSE = zeros(1,k_fold);
for i = 1:k_fold
    mvmask = missindex;
    for j = 1:k_fold
        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
    end
    mvmask(:,sum(mvmask) == 0) = (missindex(:,sum(mvmask) == 0));
    initialsample=init_nsfa(settings);
    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
    Y_hat = (finalsample.G *  finalsample.X);
    for j = 1:k_fold
        SSE(i) = SSE(i) + ...
            sum(nansum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
    end

end

RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
RMSE;

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'RMSE');
```

you need change the path `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/` and `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/` which have the NBSFA matlab code.

### creat the run.R file to run the code

In this case, we have problem in running `flashr` on RCC. so we just try other methods first.

```{}
source("../wrapper.R")
library(R.matlab)
## run the code
Y_centered = readMat("../example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)

PN_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PN",Kmax = 40)
flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash",Kmax = 40)
flashGwn_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_wn",Kmax = 40)
flashB_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_bf",Kmax = 40)
pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PMD",Kmax = 40)
soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "softImpute",Kmax = 40)
nsfa_rmse = NSF_OCV(OCVindex)

result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")
print(sessionInfo())
```

this file name is `runOCV.R`

### creat folder to track the result

```
mkdir outlog
```

### creat .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=./outlog/arrayJob_%A_%a.out
#SBATCH --error=./outlog/arrayJob_%A_%a.err
#SBATCH --array=1-20
#SBATCH --time=26:00:00
#SBATCH --partition=mstephens
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=8000

######################
# Begin work section #
######################

# Print this sub-job's task ID
module load R
module load matlab

mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose ../runOCV.R
```

this file is name as `OCV.sbatch`

## run OCV

now in your folder, you should have:

```
[weidong@midway2-login1 Breastcancer]$ ls
OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
```

```
sbatch OCV.sbatch
```

## to get the result

```{}
T = 20
results = matrix(NA,ncol = 7, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./boxplot.rds")
```

## plot the result

```{r}
# this is the independent result

library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label, myColors){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
    geom_boxplot()+
    # geom_violin()+
    ggtitle(title) +  theme_bw()+  scale_color_manual(values=myColors)+
    theme(legend.position= legend_position, legend.text=element_text(size=15),
          plot.title = element_text(size = 15, face = "bold"),
          axis.text.y = element_text(size =12.6),
          axis.text.x = element_text(size =12.6,angle = 45, hjust = 1))
  p
}

setwd("~/HG/flash_workflow/analysis")

fill_color = c("red","yellow3", "tan1","springgreen3", "springgreen","deepskyblue3","deepskyblue")

##############  MovieLens data

ML_res = readRDS("../data/output/missingdata/MovieLens/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
# c("flash","NBSFA","PMD","softImpute")
ml_res = readRDS("../data/output/missingdata/MovieLens/boxplot_cv.rds")
# flashG_rmse,flashGwn_rmse,pmd_rmse,soft_rmse
ml_res[1:20,] = sqrt((ml_res[1:20,])^2 * (943*1682)/(100000))
ml_res = ml_res[1:20,]
ML_res = ML_res[-c(2,13,17,21,29,37,62,76,77,93,95,100),]
ML_res = ML_res[1:20,]
# add PN result
ML_pn_res = readRDS("../data/output/missingdata/MovieLens/box_cv_pn.rds")
# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(ML_pn_res[,1],ML_pn_res[,2],ML_res[,3],ml_res[,3],ML_res[,2],ML_res[,4],ml_res[,4])

x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotM = plot_res(output,"MovieLens data",x_label = x_label,myColors = fill_color)

############ Breast Cancer data

BC_res = readRDS("../data/output/missingdata/BreastCancer/BreastCancer_box.rds")
BC_res = BC_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
bc_res = readRDS("../data/output/missingdata/BreastCancer/box_Breast.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(bc_res[,2],bc_res[,1],BC_res[,3],bc_res[,6],BC_res[,2],BC_res[,4],bc_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotB = plot_res(output,"Breast cancer data",x_label = x_label,myColors = fill_color)

############ GTEx data

GZ_res = readRDS("../data/output/missingdata/GTExZsocre/gtexzscore_box.rds")
GZ_res = GZ_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
gz_res = readRDS("../data/output/missingdata/GTExZsocre/box_cv.rds")
# flashG_rmse,nsfa_rmse,pmd_rmse,soft_rmse
gz_pn_res = readRDS("../data/output/missingdata/GTExZsocre/box_pn.rds")
gz_pn_res[c(23,25,29,34),] = NA
gz_pn_res = matrix(as.numeric(gz_pn_res[21:40,]),ncol = 2)

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(gz_res[,1],gz_pn_res[,2],GZ_res[,3],gz_res[,3],GZ_res[,2],GZ_res[,4],gz_res[,4])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotG = plot_res(output,"GTEx data",x_label = x_label,myColors = fill_color)

############ Text data

PT_res = readRDS("../data/output/missingdata/DenoiseRtext/president_box.rds")
PT_res = PT_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
pt_res = readRDS("../data/output/missingdata/DenoiseRtext/box_president.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(pt_res[,3],pt_res[,1],PT_res[,3],pt_res[,6],PT_res[,2],PT_res[,4],pt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotP = plot_res(output,"Presidential address data",x_label = x_label,myColors = fill_color)

############ Tumor data

DT_res = readRDS("../data/output/missingdata/DenoiseRtumor/denoiseRtumor_box.rds")
DT_res = DT_res[1:20,]
# c("flash","NBSFA","PMD","softImpute")
dt_res = readRDS("../data/output/missingdata/DenoiseRtumor/box_denoiseTumor.rds")
# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")

# the result order would be flashG PMD PMDcv NSFA SF SFcv
output = cbind(dt_res[,2],dt_res[,1],DT_res[,3],dt_res[,6],DT_res[,2],DT_res[,4],dt_res[,7])
x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
plotT = plot_res(output,"Tumor data",x_label = x_label,myColors = fill_color)
```

```{r}
# gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))
```

## other information

please click here to check other information [Missing data OCV settings](OCVmissdataSetting.html)

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```