### 2018-02-11 analysis/index.Rmd
Files commited by wflow_commit.

---
title: "Home"
output:
  html_document:
    toc: false
---

## Apply flashr for the examples in FLASH paper

### simulations

Click here to check the [RMSE for the rank one example in the paper](MSErank1sim.html)

Click here to check the [RMSE for the rank three example in the paper](simulation3.html) and [RRMSE](MSErank3sim.html)

### GTEx Z-score

Here is a example to run `flashr` on RCC [RCC template](GTEXzscoreTemplate.html)

compare with PCA result[compare with SVD](PCAresult.html)

### missing data imputation

click here to check the template for [missing data imputation with OCV](templateforOCV.html) and [CV template](CVtemplate.html)

## Other documents

[old experiments](oldanalysis.html)

### Orthogonal CV

Click on this [orthogonal CV](miss_position.html) to see my example.